---
title: "Baseline imputation"
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: 
    html_document:
      keep_md: false
params:
  outcome_min: 28
  outcome_max: 35
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(here)
require(mice)
require(naniar)
here::i_am(file.path("hcq_pooling_analysis", "2_Baseline_imputation.Rmd"))
source(here("hcq_pooling_analysis", "common.R"))

out_stub <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir <- here("hcq_pooling_analysis", "output", out_stub)
```

Key parameter: the number of imputations:

```{r}
n_imputations <- 20
```

### Load participant data

Read patient data:

```{r}
patients <- read_rds(file.path(output_dir, "patients.rds"))
```

Select those patients who will participate in the imputation step:

```{r}
patients_for_imputation <-
  patients %>%
  filter(rand_grp != "SCREEN FAIL")

nrow(patients); nrow(patients_for_imputation)
```

These are the patients who were successfully randomized/assigned to a treatment arm:

```{r}
sum(!is.na(patients$treat))
```

### Select/create variables to be used in the imputation model

Things to do:

-   Due to extensive missingness, we're not using the HIV or AIDS variables
-   We also need to collapse smoking and vaping to never/ever binary versions
-   We need to set to NA negative values of days since symptom onset.

```{r}
comorbidities_all <-
  c("aids",
    "crbr_vasc_dis",
    "mi",
    "cng_hrt_flr",
    "dementia",
    "copd",
    "asthma",
    "hypertension",
    "hiv",
    "tumor",
    "liver_dis",
    "diabetes",
    "smoking",
    "vaping"
)

# Could move this to common.R
comorbidities_imputation <-
  c(# "aids",
    "crbr_vasc_dis",
    "mi",
    "cng_hrt_flr",
    "dementia",
    "copd",
    "asthma",
    "hypertension",
    # "hiv",
    "tumor",
    "liver_dis",
    "diabetes",
    "smoking_binary", # "smoking",
    "vaping_binary" # "vaping"
)

other_covariates_imputation <-
  c(# In the outcome model(s)—excluding
    # treatment and outcome
    "sex_fct",
    "age_5y",
    "bmi",
    "niaid_baseline_fct",
    "sym_onst_days_bfr_enrdt",
    # Others
    "race_simplified_fct",
    "ethnic_fct"
  )
```

Create the table.

Here we implement the choice to impute the following values:

-   Implicit symptom onset _after_ enrollment.
-   BMIs outside of 10 ≤ bmi ≤ 70 set to missing.

```{r}
imputation_tbl <-
  patients_for_imputation %>%
  # Create binary/logical comorbidities
  mutate(
    # as.numeric becase we want these to be 0,1 valued to match the others
    hiv_binary     = as.numeric((hiv == 0)),
    smoking_binary = as.numeric((smoking == 0)),
    vaping_binary  = as.numeric((vaping == 0))
    ) %>%
  mutate_at(comorbidities_imputation, factor) %>%
  # Remove negative symptom onset days before enrollment
  mutate(
    sym_onst_days_bfr_enrdt = 
      ifelse(sym_onst_days_bfr_enrdt < 0, NA, 
             sym_onst_days_bfr_enrdt),
    # Remove extreme BMIs
    bmi =
      ifelse((bmi >= 10) & (bmi <= 70),
	     bmi,
	     NA)
  ) %>%
  select(
    siteid, patient_id,
    !!comorbidities_imputation,
    !!other_covariates_imputation
  )

write_rds(imputation_tbl, file.path(output_dir, "imputation_tbl.rds"))
```

### Examine missingness patterns

```{r}
gg_miss_var(imputation_tbl %>% select(-siteid, patient_id))
```

```{r}
mice::md.pattern(
  imputation_tbl %>% select(-siteid, -patient_id),
  rotate.names = TRUE)
```

By site:

```{r}
# TO DO

# mice::md.pattern(
#   imputation_tbl %>% select(-siteid, -patient_id),
#   plot = FALSE) %>%
#   as.data.frame() %>%
#   rownames_to_column("count")
```

By treatment arm:

Possibly by both?

## Running MICE

Need to set a specific seed:

```{r message=FALSE}
mice_fit <- mice::mice(
  imputation_tbl %>% 
    select(-siteid, -patient_id), 
  seed = 20200524,
  maxit = 20,
  m = n_imputations # This strongly affects how long BRMS takes to fit
  )
```

## Examining MICE fit

Get a sense of the mixing:

```{r}
plot(mice_fit)
```

See which method was used for each variable:

```{r}
summary(mice_fit)
```

```{r}
stripplot(mice_fit, jitter.data = TRUE)
```

```{r}
densityplot(mice_fit)
```

```{r}
bwplot(mice_fit)
```

It would be great to produce more diagnostics for the categorical variables (comorbidities).

We might also compare the values for their sum.

Note: this is done *ignoring* siteid (rather than including it as a predictor, or doing multilevel MI).

### Write out the imputations

```{r}
write_rds(mice_fit, file.path(output_dir, "mice_fit_object.rds"))
```

## Adding response and model covariates


Now to:

-   pull out the list of imputed datasets
-   join outcome variable, treatment assignment, and azithro, which weren't used in the imputation
-   compute variables needed for the outcome model, in particular the comorbidity count

```{r}

# Could move this to common.R
comorbidities_imputation <-
  c(# "aids",
    "crbr_vasc_dis",
    "mi",
    "cng_hrt_flr",
    "dementia",
    "copd",
    "asthma",
    "hypertension",
    # "hiv",
    "tumor",
    "liver_dis",
    "diabetes",
    "smoking_binary", # "smoking",
    "vaping_binary" # "vaping"
)

add_postbaseline_and_model_vars <-
  function(d) {
      comorbidity_count <-
        rowSums((as.matrix(d[,comorbidities_imputation])) == 1 | (as.matrix(d[,comorbidities_imputation]) == TRUE))
      
      d %>% 
        bind_cols(imputation_tbl %>% select(siteid, patient_id)) %>%
        left_join(
          patients %>% select(patient_id, treat, azithro, niaid_outcome),
          by = "patient_id"
        ) %>%
        mutate(
          sex_model = ifelse(sex_fct == "female", 1/2, -1/2),
          age_model = (age_5y - 60)/10,
          bmi_model = (bmi - 25)/5,
          comorbidity_count = comorbidity_count,
          niaid_baseline_numeric_model = 
                # niaid_baseline_fct goes from 2:5, 
                # converted to 1:4 by as.numeric
            5 - (as.numeric(niaid_baseline_fct) + 1),
          treat = relevel(factor(treat), ref = "no_HCQ")
        ) %>%
        as_tibble()
    }

mice_df_list <- 
  mice::complete(mice_fit, action = "all", include = FALSE) %>%
  map(
    add_postbaseline_and_model_vars
  )
```
We also need to a table of the raw data with baseline missingness intact:

```{r}
data_tbl <- add_postbaseline_and_model_vars(
  imputation_tbl %>% select(-siteid, -patient_id))
```

Write this out:

```{r}
write_rds(data_tbl, file.path(output_dir, "data_tbl.rds"))
```

Also write out the MICE df list:

```{r}
write_rds(mice_df_list, file.path(output_dir, "mice_complete_df_list.rds"))
```

... and a `mids` object with the additional variables:

```{r}
mice_fit_withvars <- 
  mice::complete(mice_fit, action = "long", include = TRUE) %>%
  nest(-.imp) %>%
  mutate(data = map(data, add_postbaseline_and_model_vars)) %>%
  unnest(data) %>%
  as.mids()

write_rds(mice_fit_withvars, file.path(output_dir, "mice_fit_object_withvars.rds"))
```






```{r}
sessionInfo()
```



```{r}
Sys.time()
```

