---
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: false
params:
  fit_name: "main_fit_Vanderbilt_only"
  outcome_min: 28
  outcome_max: 35
title: "`r paste0('Outcome model fit', params$outcome_min, params$outcome_max, params$fit_name, sep = '-')`"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(here)
require(brms)
require(tidybayes)
require(mice)
require(cmdstanr)

here::i_am(file.path("hcq_pooling_analysis", "3a_7_Outcome_model_fit_Vanderbilt_only.Rmd"))

source(here("hcq_pooling_analysis", "common.R"))

set_cmdstan_path(here::here("hcq_pooling_analysis", ".cmdstan", "cmdstan-2.27.0"))

out_stub         <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir       <- here("hcq_pooling_analysis", "output", out_stub)
output_model_dir <- file.path(output_dir, params$fit_name)
if(!dir.exists(output_model_dir)) {
  dir.create(output_model_dir, recursive = TRUE)
}
```

```{r}
params
```

Read multiply-imputed data:

```{r}
mice_df_list_all <-
  read_rds(here(output_dir, "mice_complete_df_list.rds"))

mice_df_list_orchid_only <-
  mice_df_list_all %>%
  # FIT ONLY TO ORCHID DATA
  map(function(d) { filter(d, siteid == "ORCHID") })

# patients <-
#   read_rds(here(output_dir, "patients.rds"))

map(mice_df_list_orchid_only, nrow)
```

Specify the primary model formula:

```{r}
orchid_only_model <-
  brms::bf(
    niaid_outcome ~
        treat*(
          sex_model +
          splines::ns(age_model, 3) +
          splines::ns(bmi_model, 3) +
          splines::ns(comorbidity_count, 3) +
          niaid_baseline_numeric_model
        ) +
        # (1 + treat || siteid) + ## Exclude site random effects
        (1 + treat || niaid_baseline_fct)
  )
```

The problem we have is that there are no individuals with outcome NCOSS of 3 in the ORCHID data.

This means that the stan code output by `brm`, as well as the stan data prepared by `brm`, have only 6 levels/5 cutpoints instead of 7 levels/6 cutpoints.

The solution, it would seem, is to 

1.  manually produce stan code as for the full model, then 
2.  apply this to manually filtered ORCHID-only stan data.

Manually producing the stan code:

```{r}
(orchid_model_stancode <- 
  make_stancode(
    formula = orchid_only_model, 
    family = cumulative,
    data = mice_df_list_all[[1]],
    prior = prior(student_t(3, 0, 10), class = sd)))
```

Manually producing stan data:

```{r}
orchid_model_standata_imp1 <-
 make_standata(
    formula = orchid_only_model, 
    family = cumulative,
    data = mice_df_list_all[[1]],
    prior = prior(student_t(3, 0, 10), class = sd))
```

Note that there are 6 thresholds (correct):

```{r}
orchid_model_standata_imp1$nthres
```

... and the responses go from 1 to 7:

```{r}
orchid_model_standata_imp1$Y %>% table()
```

Whereas if I do the same for the ORCHID data only:

```{r}
orchid_model_standata_imp1_orchid_only <-
 make_standata(
    formula = orchid_only_model, 
    family = cumulative,
    data = mice_df_list_orchid_only[[1]],
    prior = prior(student_t(3, 0, 10), class = sd))
```

```{r}
orchid_model_standata_imp1_orchid_only$nthres
```

... and the responses go from 1 to 6!:

```{r}
orchid_model_standata_imp1_orchid_only$Y %>% table()
```

Cf. the raw data coding:

```{r}
mice_df_list_orchid_only[[1]]$niaid_outcome %>% as.numeric() %>% table()
```

Filtering the "all data" standata object manually:

```{r}
names(orchid_model_standata_imp1)
```

```{r}
orchid_indices_among_nonmissing_y <- 
  (mice_df_list_all[[1]] %>% filter(!is.na(niaid_outcome)))$siteid == "ORCHID" 

c(n_nonmissing = length(orchid_indices_among_nonmissing_y),
  n_nonmissing_orchid = sum(orchid_indices_among_nonmissing_y))
```

```{r}
map(orchid_model_standata_imp1, dim)
map(orchid_model_standata_imp1, length)
```

```{r}
orchid_model_standata_imp1_orchid_only_manual <- orchid_model_standata_imp1

orchid_model_standata_imp1_orchid_only_manual$N <- 
  sum(orchid_indices_among_nonmissing_y)

orchid_model_standata_imp1_orchid_only_manual$Y <- 
  orchid_model_standata_imp1$Y[orchid_indices_among_nonmissing_y]

orchid_model_standata_imp1_orchid_only_manual$X <- 
  orchid_model_standata_imp1$X[orchid_indices_among_nonmissing_y,]

orchid_model_standata_imp1_orchid_only_manual$Z_1_1 <- 
  orchid_model_standata_imp1$Z_1_1[orchid_indices_among_nonmissing_y]

orchid_model_standata_imp1_orchid_only_manual$Z_1_2 <- 
  orchid_model_standata_imp1$Z_1_2[orchid_indices_among_nonmissing_y]

orchid_model_standata_imp1_orchid_only_manual$J_1 <- 
  orchid_model_standata_imp1$J_1[orchid_indices_among_nonmissing_y]

attr(orchid_model_standata_imp1_orchid_only_manual, "class") <- "standata"
```

```{r}
map(orchid_model_standata_imp1_orchid_only_manual, dim)
map(orchid_model_standata_imp1_orchid_only_manual, length)
```

Try to fit with this:

```{r}
system.time(
  imp_1_orchid_only_fit <- rstan::stan(model_code = orchid_model_stancode, data = orchid_model_standata_imp1_orchid_only_manual))
```

Check that there are 6 cutpoints:

```{r}
bayesplot::mcmc_dens(imp_1_orchid_only_fit, regex_pars = "b_Intercept")
```

Try turning this into a `brmsfit` object:

```{r}
orchid_only_test_brmfit_1 <- brm(
    formula = orchid_only_model, 
    family = cumulative,
    data = mice_df_list_all[[1]],
    prior = prior(student_t(3, 0, 10), class = sd),
    empty = TRUE)

orchid_only_test_brmfit_1$fit <- imp_1_orchid_only_fit
orchid_only_test_brmfit_1 <- rename_pars(orchid_only_test_brmfit_1)
orchid_only_test_brmfit_1
```

Create a function for updating the "standata" to only include ORCHID:

```{r}
filter_standata_to_orchid <- function(old_standata, orchid_indices) {
  new_standata <- old_standata

  new_standata$N <- 
    sum(orchid_indices)
  
  new_standata$Y <- 
    old_standata$Y[orchid_indices]
  
  new_standata$X <- 
    old_standata$X[orchid_indices,]
  
  new_standata$Z_1_1 <- 
    old_standata$Z_1_1[orchid_indices]
  
  new_standata$Z_1_2 <- 
    old_standata$Z_1_2[orchid_indices]
  
  new_standata$J_1 <- 
    old_standata$J_1[orchid_indices]
  
  attr(new_standata, "class") <- "standata"

  return(new_standata)
  }
```

Create "standata" objects for each of the imputations:

```{r}
mice_standata_list_orchid <- 
  map(mice_df_list_all, function (d) {
    make_standata(
      formula = orchid_only_model, 
      family = cumulative,
      data = d,
      prior = prior(student_t(3, 0, 10), class = sd)) %>%
    filter_standata_to_orchid(orchid_indices_among_nonmissing_y)
  })
```

Model fitting function:

```{r}
ncores <- parallel::detectCores() - 1

fit_orchid_standata <- function(orchid_standata) {
  rstan::stan(
    model_code = orchid_model_stancode, 
    data = orchid_standata,
    cores = ncores,
    iter = 3000,
    thin = 3,
    control = list(adapt_delta = 0.999),
    seed = 20200524)
}
```

Fit each of these models:

```{r}
stan_fit_list_orchid <-
  map(mice_standata_list_orchid, fit_orchid_standata)
```

Create `brmfit` objects:

```{r}
brmfit_list <-
  map2(stan_fit_list_orchid, mice_df_list_all,
  function(stanfit, full_df) {
    brm_fit <- brm(
      formula = orchid_only_model, 
      family = cumulative,
      data = mice_df_list_all[[1]],
      prior = prior(student_t(3, 0, 10), class = sd),
      empty = TRUE)

    brm_fit$fit <- stanfit
    brm_fit <- rename_pars(brm_fit)
    return(brm_fit)
  }
  )
```

Now combine these:

```{r}
brmfit_combined <- combine_models(mlist = brmfit_list, check_data = FALSE)

# The source for brm_multiple does this, and we make use of it
brmfit_combined$rhats <- do.call(rbind, map(brmfit_list, rhat))
```

Fit the primary outcome model:

```{r message=FALSE}
write_rds(brmfit_combined, file.path(output_model_dir, paste0(params$fit_name, ".rds")))
```


```{r}
sessionInfo()
```



```{r}
Sys.time()
```

