---
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: 
    html_document:
      toc: true
      keep_md: false
params:
  fit_name: "main_fit"
  outcome_min: 28
  outcome_max: 35
title: "`r paste0('Subgroup effects - tables and plots -', params$outcome_min, params$outcome_max, params$fit_name, sep = '-')`"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(here)
here::i_am(file.path("hcq_pooling_analysis", "4b_Subgroup_effect_estimates.Rmd"))
source(here("hcq_pooling_analysis", "common.R"))

require(brms)
require(tidybayes)

require(mice)

require(bayesplot)
bayesplot_theme_set(theme_cowplot())
bayesplot_theme_update(
  strip.background = element_blank(),
  strip.text.y = element_text(angle = 0),
  strip.text = element_text(face = "bold")
)

require(MASS, exclude = "select") # clashes with dplyr

out_stub         <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir       <- here("hcq_pooling_analysis", "output", out_stub)
output_model_dir <- file.path(output_dir, params$fit_name)
fit_file         <- file.path(output_model_dir, paste0(params$fit_name, ".rds"))

subgroup_dir <- file.path(output_model_dir, "subgroup_effects")
```

## Loading model and data

```{r}
brm_fit <- read_rds(fit_file)
data_tbl <- read_rds(file.path(output_dir, "data_tbl.rds"))
data_tbl_annotated <- read_rds(file.path(output_model_dir, "data_tbl_annotated.rds"))
```

Load posterior fitted/predictive tables:

```{r}
data_tbl_po_pp              <- read_rds(file.path(subgroup_dir,
                                                  paste0(params$fit_name, "_data_tbl_po_pp.rds")))
data_tbl_po_pp_matched_pair <- read_rds(file.path(subgroup_dir,
                                                  paste0(params$fit_name, "_data_tbl_po_pp_matched_pair.rds")))
data_tbl_po_fitted          <- read_rds(file.path(subgroup_dir,
                                                  paste0(params$fit_name, "_data_tbl_po_fitted_.rds")))
```

## Functions for the different estimands

We want a function to compute the "proportional odds estimand" using `polr`. If there are only two levels' worth of responses in the dataset, then use a logistic regression.

```{r}
polr_tidy_fit <- function(treat, outcome, ...) {
  polr(outcome ~ treat,
       Hess = TRUE,
       ...
       ) %>%
    broom::tidy(conf.int = TRUE, conf.level = .95) %>%
    filter(term == "treatHCQ") %>% 
    transmute(
      logor_hat      = estimate, 
      logor_95_lower = conf.low, 
      logor_95_upper = conf.high, 
      logor_se       = std.error)
}

polr_est_only <- function(treat, outcome, ...) {
  logor_hat <-
    polr(outcome ~ treat,
         Hess = TRUE,
         ...
         ) %>%
    coef()
  
  return(logor_hat)
}

logit_tidy_fit <- function(treat, outcome, ...) {
  glm(outcome ~ treat,
      family = binomial(),
      ...
      ) %>%
    broom::tidy(conf.int = TRUE, conf.level = .95) %>%
    filter(term == "treatHCQ") %>% 
    transmute(
      logor_hat      = estimate, 
      logor_95_lower = conf.low, 
      logor_95_upper = conf.high, 
      logor_se       = std.error)
}

logit_est_only <- function(treat, outcome, ...) {
  logor_hat <-
    (glm(outcome ~ treat,
      family = binomial(),
      ...
      ) %>%
    coef())[2]
  
  return(logor_hat)
}

rd_plugin <- function(treat, outcome) {
  # Exclude from the numerator and denominator those with missing outcomes
  treat   <- treat[!is.na(outcome)]
  outcome <- outcome[!is.na(outcome)]
  
  pt_obj <- prop.test(

    c(sum(treat & outcome), sum((!treat) & outcome)),
    c(sum(treat)          , sum((!treat))          ), 
    conf.level = .95)
  
  est <- diff(rev(pt_obj$estimate))
  names(est) <- "diff"
  
  return(c(est = est,
           lower_95 = pt_obj$conf.int[1],
           upper_95 = pt_obj$conf.int[2]))
}

prop_plugin <- function(treat_grp, outcome) {
  treat_grp <- treat_grp[!is.na(outcome)]
  outcome   <- outcome[!is.na(outcome)]
  
    pt_obj <- prop.test(
      # Exclude from the numerator and denominator those with missing outcomes
      sum(treat_grp & outcome),
      sum(treat_grp),
      conf.level = .95)
  
  est <- pt_obj$estimate
  names(est) <- "prop"
  
  return(c(est = est,
           lower_95 = pt_obj$conf.int[1],
           upper_95 = pt_obj$conf.int[2]))
}
```

There's no need to pool over multiple imptuations here, since we're not making use of any of the baseline covariates.

Check the correlation between the mean of cumulative log odds ratios and the `polr` estimand:

```{r}
# Find somewhere for this ...

# qplot(mean_logor, logor_hat,
#       data = polr_by_age_posterior_superpop,
#       size = I(.3)) +
#   geom_abline(color = "orange") +
#   coord_fixed() +
#   ylab("polr estimand") +
#   geom_label(x = 1, y = -1, label = 
#                str_c("corr = ",
#                  cor(polr_by_age_posterior_superpop$mean_logor,
#                      polr_by_age_posterior_superpop$logor_hat) %>%
#                    round(2)))
```

## Functions to compute subgroup effects from tables of posterior samples

Create functions to perform the standardized analysis for any given factor:

### Finite-sample, matched estimand (not included)

```{r}
grouped_matched_posterior <- 
  function(paired_pp_draws, grouping_factor) {
    paired_pp_draws %>%
    mutate(
      .prediction = case_when(
        assignment == "observed"       ~ as.numeric(niaid_outcome),
        assignment == "counterfactual" ~ as.numeric(.prediction)) %>%
          factor(levels = 1:7, labels = niaid_levels, ordered = TRUE)) %>%
    group_by({{ grouping_factor }}, .draw) %>%
    summarize(
      logor_hat =
        possibly(
          polr_est_only, 
          otherwise = NA, 
          quiet = FALSE)(
        treat, .prediction)) %>%
    ungroup()
  }
```

Summary functions:

```{r}
summarize_grouped_matched_posterior <-
  function(grouped_posterior, grouping_factor) {
    grouped_posterior %>%
      ungroup() %>% 
      group_by({{ grouping_factor }}) %>%
    summarise(
      logor_95_lower    = quantile(logor_hat, .025, na.rm = TRUE),
      logor_95_upper    = quantile(logor_hat, .975, na.rm = TRUE),
      logor_se          = sd(logor_hat, na.rm = TRUE),
      prop_na_logor_hat = mean(is.na(logor_hat)),
      logor_hat         = median(logor_hat, na.rm = TRUE) # this needs need to be last!
    )
  }

```

### Superpopulation estimand

```{r}
grouped_superpop_posterior <-
  function(paired_fitted_draws, grouping_factor) {
    paired_fitted_draws %>%
    mutate(.category = as.numeric(substr(.category, 1, 2))) %>%
    group_by(.draw, treat, .category, {{grouping_factor}}) %>%
    # Compute expected proportions within each treatment x group x category stratum
    summarise(ave_prob = mean(value)) %>%
    # Compute (logit) eCDFs within treatment x group stratum
    group_by(.draw, treat, {{grouping_factor}}) %>%
    arrange(.category) %>%
    mutate(cuml_prob = cumsum(ave_prob),
           cuml_logit = qlogis(cuml_prob)) %>%
    # Compute estimands of interest within each group stratum
    group_by(.draw, {{grouping_factor}}) %>%
    arrange(treat, .category) %>%
    summarise(
      mean_logor = -mean(cuml_logit[treat == "HCQ"][1:6] - cuml_logit[treat == "no_HCQ"][1:6]),
      logor_hat  = 
        possibly(
          polr_est_only, otherwise = NA, quiet = FALSE)(
          treat, factor(
            .category, ordered = TRUE),
            # ensure at least 1000 in each category
            weights = round(1000 * ave_prob/min(c(ave_prob, 1 - ave_prob)))),
      
      # Survival
      est_p_surv_HCQ    = 1 - ave_prob[(.category == 1) & (treat == "HCQ")],
      est_p_surv_no_HCQ = 1 - ave_prob[(.category == 1) & (treat == "no_HCQ")],
      or_surv = (est_p_surv_HCQ/(1 - est_p_surv_HCQ))/(est_p_surv_no_HCQ/(1 - est_p_surv_no_HCQ)),
      logor_surv = log(or_surv),
      rd_surv = est_p_surv_HCQ - est_p_surv_no_HCQ,
      
      # Not (mechanical ventilation or death)
      est_p_n12_HCQ    = 1 - ave_prob[(.category <= 2) & (treat == "HCQ")],
      est_p_n12_no_HCQ = 1 - ave_prob[(.category <= 2) & (treat == "no_HCQ")],
      or_n12 = (est_p_n12_HCQ/(1 - est_p_n12_HCQ))/(est_p_n12_no_HCQ/(1 - est_p_n12_no_HCQ)),
      logor_n12 = log(or_n12),
      rd_n12 = est_p_n12_HCQ - est_p_n12_no_HCQ
      )
  }
```

Summary function:

```{r}
summarize_grouped_superpop_posterior <-
  function(grouped_posterior, grouping_factor) {
    grouped_posterior %>%
      ungroup() %>% 
      group_by({{ grouping_factor }}) %>%
    summarise(
      logor_95_lower    = quantile(logor_hat, .025, na.rm = TRUE),
      logor_95_upper    = quantile(logor_hat, .975, na.rm = TRUE),
      logor_se          = sd(logor_hat, na.rm = TRUE),
      prop_na_logor_hat = mean(is.na(logor_hat)),
      # These must go last
      logor_hat         = median(logor_hat, na.rm = TRUE),
      
      or_surv_95_lower  = quantile(or_surv, .025, na.rm = TRUE),
      or_surv_95_upper  = quantile(or_surv, .975, na.rm = TRUE),
      prop_na_or_surv   = mean(is.na(or_surv)),
      or_surv           = median(or_surv, na.rm = TRUE),
      
      or_n12_95_lower   = quantile(or_n12, .025, na.rm = TRUE),
      or_n12_95_upper   = quantile(or_n12, .975, na.rm = TRUE),
      prop_na_or_n12    = mean(is.na(or_n12)),
      or_n12            = median(or_n12, na.rm = TRUE),
      
      rd_surv_95_lower  = quantile(rd_surv, .025),
      rd_surv_95_upper  = quantile(rd_surv, .975),
      rd_surv           = median(rd_surv),
      
      rd_n12_95_lower   = quantile(rd_n12, .025),
      rd_n12_95_upper   = quantile(rd_n12, .975),
      rd_n12            = median(rd_n12),
      
      est_p_surv_HCQ_95_lower   = quantile(est_p_surv_HCQ, .025),
      est_p_surv_HCQ_95_upper   = quantile(est_p_surv_HCQ, .975),
      est_p_surv_HCQ            = median(est_p_surv_HCQ),
      
      est_p_surv_no_HCQ_95_lower   = quantile(est_p_surv_no_HCQ, .025),
      est_p_surv_no_HCQ_95_upper   = quantile(est_p_surv_no_HCQ, .975),
      est_p_surv_no_HCQ            = median(est_p_surv_no_HCQ),
      
      est_p_n12_HCQ_95_lower   = quantile(est_p_n12_HCQ, .025),
      est_p_n12_HCQ_95_upper   = quantile(est_p_n12_HCQ, .975),
      est_p_n12_HCQ            = median(est_p_n12_HCQ),
      
      est_p_n12_no_HCQ_95_lower   = quantile(est_p_n12_no_HCQ, .025),
      est_p_n12_no_HCQ_95_upper   = quantile(est_p_n12_no_HCQ, .975),
      est_p_n12_no_HCQ            = median(est_p_n12_no_HCQ)
    )
  }
```

### Simple plugin estimator

```{r}
or_fit_surv <- function(treat, outcome) {
  logit_tidy_fit(treat,  as.numeric(outcome) != 1) %>%
        transmute(
          or_surv          = exp(logor_hat), 
          or_surv_95_lower = exp(logor_95_lower), 
          or_surv_95_upper = exp(logor_95_upper))
}

or_fit_n12 <- function(treat, outcome) {
  logit_tidy_fit(treat, as.numeric(outcome) > 2) %>%
        transmute(
          or_n12 = exp(logor_hat), 
          or_n12_95_lower = exp(logor_95_lower), 
          or_n12_95_upper = exp(logor_95_upper))
}

rd_fit_surv <- function(treat, outcome) {
  fit <- rd_plugin(treat == "HCQ", as.numeric(outcome) != 1)
  
  tibble(rd_surv = fit["est.diff"],
         rd_surv_95_lower = fit["lower_95"],
         rd_surv_95_upper = fit["upper_95"])
}

rd_fit_n12 <- function(treat, outcome) {
  fit <- rd_plugin(treat == "HCQ", as.numeric(outcome) > 2)
  
  tibble(rd_n12 = fit["est.diff"],
         rd_n12_95_lower = fit["lower_95"],
         rd_n12_95_upper = fit["upper_95"])
}

fit_p_tidy <- function(treat_grp, outcome) {
  fit <- prop_plugin(treat_grp, outcome)
  
  tibble(p_hat = fit["est.prop"],
         lower_95 = fit["lower_95"],
         upper_95 = fit["upper_95"])
}
      

polr_grouped_plugin <-
  function(data_table, grouping_factor) {
    data_table %>%
    group_by({{ grouping_factor }}) %>%
    summarise(
      polr_est = 
        # Plug-in polr estimator doesn't make much sense for small n
        # or when there aren't both treatment groups represented
        list(possibly(polr_tidy_fit, otherwise = tibble(), quiet = FALSE)
          (treat, niaid_outcome)),
      or_surv_analysis =
        list(possibly(or_fit_surv, otherwise = tibble(), quiet = FALSE)
          (treat, niaid_outcome)),
      or_n12_analysis =
        list(possibly(or_fit_n12, otherwise = tibble(), quiet = FALSE)
          (treat, niaid_outcome)),
      rd_surv = list(possibly(rd_fit_surv, otherwise = tibble(), quiet = FALSE)
                     (treat, niaid_outcome)),
      rd_n12  = list(possibly(rd_fit_n12, otherwise = tibble(), quiet = FALSE)
                     (treat, niaid_outcome)),
      est_p_surv_HCQ    = list(possibly(fit_p_tidy, otherwise = tibble(p_hat = NA, lower_95 = NA, upper_95 = NA), quiet = FALSE)
                     (treat == "HCQ", as.numeric(niaid_outcome) != 1) %>%
        rename(est_p_surv_HCQ = p_hat,
               est_p_surv_HCQ_95_lower = lower_95,
               est_p_surv_HCQ_95_upper = upper_95)),
      est_p_surv_no_HCQ = list(possibly(fit_p_tidy, otherwise = tibble(p_hat = NA, lower_95 = NA, upper_95 = NA), quiet = FALSE)
                     (treat == "no_HCQ", as.numeric(niaid_outcome) != 1) %>%
        rename(est_p_surv_no_HCQ = p_hat,
               est_p_surv_no_HCQ_95_lower = lower_95,
               est_p_surv_no_HCQ_95_upper = upper_95)),
      est_p_n12_HCQ     = list(possibly(fit_p_tidy, otherwise = tibble(p_hat = NA, lower_95 = NA, upper_95 = NA), quiet = FALSE)
                     (treat == "HCQ", as.numeric(niaid_outcome) > 2) %>%
        rename(est_p_n12_HCQ = p_hat,
               est_p_n12_HCQ_95_lower = lower_95,
               est_p_n12_HCQ_95_upper = upper_95)),
      est_p_n12_no_HCQ  = list(possibly(fit_p_tidy, otherwise = tibble(p_hat = NA, lower_95 = NA, upper_95 = NA), quiet = FALSE)
                     (treat == "no_HCQ", as.numeric(niaid_outcome) > 2) %>%
        rename(est_p_n12_no_HCQ = p_hat,
               est_p_n12_no_HCQ_95_lower = lower_95,
               est_p_n12_no_HCQ_95_upper = upper_95))
      ) %>%
    unnest(c(
      polr_est, or_surv_analysis, or_n12_analysis,
      rd_surv, rd_n12, 
      est_p_surv_HCQ, est_p_surv_no_HCQ, est_p_n12_HCQ, est_p_n12_no_HCQ), 
      keep_empty = TRUE)
  }
```




```{r}
standardized_effect_analysis <- 
  function(paired_posterior_draws, fitted_posterior_draws, raw_data, grouping_factor) {
    
    # Standardize over actual (i.e., possibly missing) rather than imputed
    # covariate values
    fitted_posterior_draws <-
      left_join(fitted_posterior_draws %>% select(-{{grouping_factor}}), 
                raw_data %>% select(patient_id, {{grouping_factor}}))
    
    paired_posterior_draws <-
      left_join(paired_posterior_draws %>% select(-{{grouping_factor}}),
                raw_data %>% select(patient_id, {{grouping_factor}}))
    
    superpop_posterior <- grouped_superpop_posterior(fitted_posterior_draws, {{grouping_factor}})
      
    matched_posterior <- grouped_matched_posterior(paired_posterior_draws, {{grouping_factor}})
    
    plug_in_summary <- polr_grouped_plugin(raw_data, {{grouping_factor}})
    
    # summary table
    summary_table <- bind_rows(
      summarize_grouped_superpop_posterior(superpop_posterior, {{grouping_factor}}) %>% 
        mutate(analysis = "superpop_posterior"),
      summarize_grouped_matched_posterior(matched_posterior, {{grouping_factor}}) %>%
        mutate(analysis = "matched_posterior"),
      plug_in_summary %>% 
        mutate(analysis = "maximum likelihood")
    )
        
    return(list(
      # superpop_posterior = superpop_posterior, # -- full posterior -- in case we want more detailed plots
      # matched_posterior = matched_posterior,   # -- full posterior -- in case we want more detailed plots
      summary_table = summary_table,
      # plot = effect_summary_plot(summary_table, {{ grouping_factor }}), # this creates a huge object for some reason
      factor_count = raw_data %>% count({{ grouping_factor }}),
      factor_count_treat = raw_data %>% count({{ grouping_factor }}, treat)))
  }
```

### Plotting functions

Plotting functions:

```{r}
# basic summary plot
effect_summary_plot <- function(summary_table, grouping_factor) {
      summary_table %>%
      ggplot(aes({{ grouping_factor }}, logor_hat), data = .) +
      theme(panel.background = element_rect(fill = "grey95")) +
      background_grid(major = "y", color.major = "white", size.major = .6) +
      geom_pointrange(
        aes(ymin = logor_95_lower, ymax = logor_95_upper, color = analysis),
        position = position_dodge(width = .5))
}
```

Adding size based on n in group:

```{r}
effect_summary_plot_sized_dots <- function(
  summary_table, count_table, 
  grouping_factor,
  max_point_size = 5) {
      plot_data <- 
        summary_table %>%
        left_join(count_table) %>%
        mutate(scale_point = ifelse(analysis == "maximum likelihood", TRUE, FALSE))
      
      plot_data %>%
        ggplot(aes({{ grouping_factor }}, logor_hat), data = .) +
        theme(panel.background = element_rect(fill = "grey95")) +
        background_grid(major = "y", color.major = "white", size.major = .6) +
        geom_linerange(aes(ymin = logor_95_lower, ymax = logor_95_upper, color = analysis),
                       position = position_dodge(width = .7)) +
        geom_point(
          aes(color = analysis, alpha = scale_point,
              size = n),
          position = position_dodge(width = .7),
              shape = 21,
              fill = "grey95",
          data = plot_data) +
        geom_point(
          aes(color = analysis, alpha = I(!scale_point)),
          position = position_dodge(width = .7),
          size = 2,
          data = plot_data) +
        scale_size_area(max_size = max_point_size) +
        scale_alpha_manual(values = c(0, 1)) +
        guides(alpha = FALSE)
}
#effect_summary_plot_sized_dots(age_analysis$summary_table, data_tbl %>% mutate(age_fct = mk_age_fct(age_5y)), age_fct)
```

## Perform subgroup analyses by factor

### Age

```{r}
age_subgroups <-
  c("≤29", "30-49", "50-69", "70-79", "80+")

mk_age_fct <- function(age) {
  case_when(
    age <  30            ~ "≤29",
    age >= 30 & age < 50 ~ "30-49",
    age >= 50 & age < 70 ~ "50-69",
    age >= 70 & age < 79 ~ "70-79",
    age >= 80            ~ "80+"
    ) %>%
    factor(age_subgroups,
           ordered = TRUE)
}
```

```{r}
system.time(age_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
    mutate(age_fct = mk_age_fct(age_5y)),
  fitted_posterior_draws = data_tbl_po_fitted %>%
    mutate(age_fct = mk_age_fct(age_5y)),
  raw_data = data_tbl %>%
    mutate(age_fct = mk_age_fct(age_5y)), 
  age_fct))
```

```{r}
write_rds(age_analysis, file.path(subgroup_dir, paste0(params$fit_name, "_age_analysis.rds")))
```

### Study

```{r}
site_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp,
  fitted_posterior_draws = data_tbl_po_fitted,
  raw_data = data_tbl, 
  siteid)
```

```{r}
write_rds(site_analysis, file.path(subgroup_dir, paste0(params$fit_name, "_site_analysis.rds")))
```

### BMI

```{r}
bmi_subgroups <-
  c("≤20", "20-25", "25-30", "30-35", ">35")

mk_bmi_fct <- function(bmi) {
  case_when(
    bmi <= 20             ~ "≤20",
    bmi >  20 & bmi <= 25 ~ "20-25",
    bmi >  25 & bmi <= 30 ~ "25-30",
    bmi >  30 & bmi <= 35 ~ "30-35",
    bmi >  35             ~ ">35"
    ) %>%
    factor(levels = bmi_subgroups,
           ordered = TRUE)
}
```

```{r}
bmi_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
    mutate(bmi_fct = mk_bmi_fct(bmi)),
  fitted_posterior_draws = data_tbl_po_fitted %>%
    mutate(bmi_fct = mk_bmi_fct(bmi)),
  raw_data = data_tbl %>%
    mutate(bmi_fct = mk_bmi_fct(bmi)), 
  bmi_fct)
```

```{r}
write_rds(bmi_analysis, file.path(subgroup_dir, paste0(params$fit_name, "bmi_analysis.rds")))
```

### Sex

```{r}
sex_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp,
  fitted_posterior_draws = data_tbl_po_fitted,
  raw_data = data_tbl,
  sex_fct)
```

```{r}
write_rds(sex_analysis, file.path(subgroup_dir, paste0(params$fit_name, "sex_analysis.rds")))
```

```{r}
effect_summary_plot(sex_analysis$summary_table, sex_fct)
```

### Baseline NCOSS

```{r}
baseline_niaid_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp,
  fitted_posterior_draws = data_tbl_po_fitted,
  raw_data = data_tbl,
  niaid_baseline_fct)
```

```{r}
write_rds(baseline_niaid_analysis, file.path(subgroup_dir, paste0(params$fit_name, "baseline_niaid_analysis.rds")))
```

```{r}
effect_summary_plot(baseline_niaid_analysis$summary_table, niaid_baseline_fct)
```

### Baseline risk

```{r}
data_tbl_annotated <-
  data_tbl_annotated %>% 
  mutate(
    risk_score_quintile_fct = 
      factor(cut_number(
        -lin_pred_control # Higher values correspond to *better* predicted outcome
        , n = 5, labels = FALSE)))

patient_risk_tbl <- data_tbl_annotated %>% select(patient_id, risk_score_quintile_fct)
```

```{r}
risk_score_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
   left_join(patient_risk_tbl),
  fitted_posterior_draws = data_tbl_po_fitted %>%
   left_join(patient_risk_tbl),
  raw_data = data_tbl_annotated,
  risk_score_quintile_fct
)
```

```{r}
effect_summary_plot(risk_score_analysis$summary_table, risk_score_quintile_fct)
```

```{r}
write_rds(risk_score_analysis, file.path(subgroup_dir, paste0(params$fit_name, "risk_score_analysis.rds")))
```

### Number of baseline comorbidities

```{r}
comorbidity_subgroups <-
  c("0", "1", "2", "3", "≥4")

mk_comorbidity_fct <- function(comorbidity_count) {
  case_when(
    comorbidity_count == 0 ~ "0",
    comorbidity_count == 1 ~ "1",
    comorbidity_count == 2 ~ "2",
    comorbidity_count == 3 ~ "3",
    comorbidity_count >= 4 ~ "≥4",
    ) %>%
    factor(levels = comorbidity_subgroups,
           ordered = TRUE)
}
```

```{r}
comorbidity_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
    mutate(comorbidity_fct = mk_comorbidity_fct(comorbidity_count)),
  fitted_posterior_draws = data_tbl_po_fitted %>%
    mutate(comorbidity_fct = mk_comorbidity_fct(comorbidity_count)),
  raw_data = data_tbl %>%
    mutate(comorbidity_fct = mk_comorbidity_fct(comorbidity_count)),
  comorbidity_fct)
```

```{r}
write_rds(comorbidity_analysis, file.path(subgroup_dir, paste0(params$fit_name, "comorbidities_analysis.rds")))
```

### Overall

```{r}
overall_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
    mutate(overall_fct = "overall"),
  fitted_posterior_draws = data_tbl_po_fitted %>%
    mutate(overall_fct = "overall"),
  raw_data = data_tbl %>%
    mutate(overall_fct = "overall"),
  overall_fct)
```

```{r}
effect_summary_plot(overall_analysis$summary_table, overall_fct)
```

```{r}
write_rds(overall_analysis, file.path(subgroup_dir, paste0(params$fit_name, "overall_analysis.rds")))
```

### (Post-hoc) Symptom onset days before enrollment

```{r}
data_tbl_annotated <-
  data_tbl_annotated %>% 
  mutate(
    sym_onset_fct = 
      factor(cut_number(
        sym_onst_days_bfr_enrdt, 
        n = 3),
        ordered = TRUE))

patient_sym_onset_tbl <- data_tbl_annotated %>% select(patient_id, sym_onset_fct)
```

```{r}
sym_onset_analysis <- standardized_effect_analysis(
  paired_posterior_draws = data_tbl_po_pp %>%
   left_join(patient_sym_onset_tbl),
  fitted_posterior_draws = data_tbl_po_fitted %>%
   left_join(patient_sym_onset_tbl),
  raw_data = data_tbl_annotated,
  sym_onset_fct
)
```

```{r}
effect_summary_plot(sym_onset_analysis$summary_table, sym_onset_fct)
```

```{r}
write_rds(sym_onset_analysis, file.path(subgroup_dir, paste0(params$fit_name, "sym_onset_analysis.rds")))
```


## Summarizing all this

Write out subgroup analysis table:

```{r fig.height = 10, fig.width = 8}
all_effect_summaries <-
  bind_rows(
    overall_analysis$summary_table        %>% mutate(subgroups = "overall", level = overall_fct),
    site_analysis$summary_table           %>% mutate(subgroups = "site",    level = siteid),
    sex_analysis$summary_table            %>% mutate(subgroups = "sex",     level = sex_fct),
    age_analysis$summary_table            %>% mutate(subgroups = "age",     level = age_fct),
    bmi_analysis$summary_table            %>% mutate(subgroups = "bmi",     level = bmi_fct),
    comorbidity_analysis$summary_table    %>% mutate(subgroups = "# comorbidities", level = comorbidity_fct),
    baseline_niaid_analysis$summary_table %>% mutate(subgroups = "baseline_niaid", level = niaid_baseline_fct),
    risk_score_analysis$summary_table     %>% mutate(subgroups = "baseline_risk_quintile", level = scales::label_ordinal()(as.numeric(risk_score_quintile_fct))),
    sym_onset_analysis$summary_table      %>% mutate(subgroups = "sym_onset_fct", level = sym_onset_fct)
  ) %>%
  select(-c(overall_fct, siteid, sex_fct, age_fct, bmi_fct, comorbidity_fct, niaid_baseline_fct, risk_score_quintile_fct, -sym_onset_fct))

# level_order <- unique(all_effect_summaries$level)[
#   str_extract(unique(all_effect_summaries$level))
# ]
```

```{r}
# Need to mutate(siteid = level, subgroups = "site") & analogously
all_group_sizes <-
  bind_rows(
    data_tbl %>% 
      mutate(overall_fct = "overall") %>% 
      count(overall_fct) %>%
      mutate(subgroups = "overall", level = overall_fct) %>%
      select(-overall_fct),
    data_tbl %>% count(siteid) %>% mutate(subgroups = "site",    level = siteid) %>% select(-siteid),
    data_tbl %>% count(sex_fct) %>% mutate(subgroups = "sex",     level = sex_fct) %>% select(-sex_fct),
    data_tbl %>% mutate(age_fct = mk_age_fct(age_5y)) %>% count(age_fct) %>% mutate(subgroups = "age",     level = age_fct) %>% select(-age_fct),
    data_tbl %>% mutate(bmi_fct = mk_bmi_fct(bmi)) %>% count(bmi_fct)%>% mutate(subgroups = "bmi",     level = bmi_fct) %>% select(-bmi_fct),
    data_tbl %>% count(niaid_baseline_fct) %>% mutate(subgroups = "baseline_niaid", level = niaid_baseline_fct) %>% select(-niaid_baseline_fct),
    data_tbl %>% mutate(comorbidity_fct = mk_comorbidity_fct(comorbidity_count)) %>% count(comorbidity_fct) %>% mutate(subgroups = "# comorbidities", level = comorbidity_fct) %>% select(-comorbidity_fct),
    data_tbl_annotated %>% count(risk_score_quintile_fct) %>% mutate(subgroups = "baseline_risk_quintile", level = scales::label_ordinal()(as.numeric(risk_score_quintile_fct))) %>% select(-risk_score_quintile_fct),
    data_tbl_annotated %>% count(sym_onset_fct) %>% mutate(subgroups = "sym_onset_fct", level = sym_onset_fct)
  )

write_tsv(all_effect_summaries %>% left_join(all_group_sizes),
          file.path(subgroup_dir, "subgroup_analysis_table.tsv"))
```

## Plots (Figure 1)

An order for the different kinds of subgroups, and their levels:

```{r}
grouping_order <-
  c("overall",
    "site",
    "baseline_niaid",
    "age",
    "sex",
    "# comorbidities",
    "bmi",
    "sym_onset_fct",
    "baseline_risk_quintile"
    )

level_order <-
  c(bmi_subgroups,
    age_subgroups,
    comorbidity_subgroups,
    levels(data_tbl$niaid_baseline_fct),
    as.character(arrange(count(data_tbl, siteid), desc(n))$siteid), # sites in descending n order
    c("female", "male"),
    "overall",
    scales::label_ordinal()(1:5),
    levels(patient_sym_onset_tbl$sym_onset_fct),
    NA, "(Missing)"
    )

all_group_sizes <- all_group_sizes %>%
  mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
         level = factor(level, levels = rev(level_order), ordered = TRUE))
```

Produce an overall summary plot:

```{r fig.width=9, fig.height=12}
all_effect_summaries %>%
  mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
         level = factor(level, levels = rev(level_order), ordered = TRUE)) %>%
  effect_summary_plot_sized_dots(
  all_group_sizes, level) +
  coord_flip(ylim = c(-4, 2)) +
  facet_grid(subgroups~., space = "free_y", scales = "free_y") +
  # theme(strip.placement = "outside", strip.text = element_text(angle = 0)) +
  background_grid(major = "x", color.major = "white")
```

The same, but without the finite sample/matched posterior estimates, and filtering out NAs:

```{r fig.width=9, fig.height=10}
odds_breaks <- c(.01, .1, .2, .5, 1, 2, 5, 10)

fig2_data <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE))

(figure_2_subgroups <- 
   fig2_data %>% 
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = log(c(.1, 10)) # HARD CODING THIS
              ) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous(
    labels = str_c(odds_breaks, "×"), 
    breaks = log(odds_breaks)) + 
    ylab("proportional odds ratio")
    )

ggsave(filename = file.path(subgroup_dir, paste0("figure_2.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_2.svg")), width = 9, height = 10)
```

Corresponding table:

```{r}
formatted_effect_table <-
  all_effect_summaries %>% 
  mutate(
    estimate_polr = ci_fmt(
      exp(logor_hat),
      exp(logor_95_lower),
      exp(logor_95_upper),
      trim = TRUE),
    estimate_rd_surv = ci_fmt(
      rd_surv,
      rd_surv_95_lower,
      rd_surv_95_upper,
      trim = TRUE)) %>% 
  select(analysis, subgroups, level, estimate_polr, estimate_rd_surv) %>% 
  filter(
    analysis %in% c("maximum likelihood", "superpop_posterior")) %>% 
  pivot_wider(names_from = analysis, values_from = c(estimate_polr, estimate_rd_surv)) %>% 
  mutate(subgroups = factor(subgroups, 
                            levels = grouping_order,
                            ordered = TRUE),
         level = factor(level, 
                        levels = rev(level_order), 
                        ordered = TRUE)) %>% 
  arrange(subgroups, rev(level))

write_tsv(formatted_effect_table, file.path(subgroup_dir, paste0("formatted_effect_table.tsv")))

formatted_effect_table %>% knitr::kable()
```

Absolute risk and risk differences by subgroup:

```{r fig.width=9, fig.height=10}
rd_breaks <- seq(-1, 1, .2)

(fig_rd <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE)) %>% 
  # HACK! #############
  mutate(logor_hat = rd_surv, logor_95_lower = rd_surv_95_lower, logor_95_upper = rd_surv_95_upper) %>%
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = c(-.3, .3) # HARD CODING THIS
              ) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous(
    breaks = rd_breaks,
    labels = scales::percent) + 
    ylab("risk difference for survival") +
    ggtitle("Estimated risk differences of survival (HC/CQ versus control): pooled and subgroup analyses"))

ggsave(filename = file.path(subgroup_dir, paste0("figure_rd.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_rd.svg")), width = 9, height = 10)
```


```{r fig.width=9, fig.height=10}
risk_breaks <- seq(0, 1, .2)

(fig_p_surv_HCQ <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE)) %>% 
  # HACK! #############
  mutate(logor_hat = est_p_surv_HCQ, 
         logor_95_lower = est_p_surv_HCQ_95_lower, 
         logor_95_upper = est_p_surv_HCQ_95_upper) %>%
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = c(0, 1)) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous( 
    breaks = risk_breaks,
    labels = scales::percent_format(accuracy = 1)) + 
    ylab("survival rate under HC/CQ") +
    ggtitle("Estimated survival rate under HC/CQ: pooled and subgroup analyses"))

ggsave(filename = file.path(subgroup_dir, paste0("figure_p_surv_HCQ.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_p_surv_HCQ.svg")), width = 9, height = 10)
```

```{r fig.width=9, fig.height=10}
risk_breaks <- seq(0, 1, .2)

(fig_p_surv_no_HCQ <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE)) %>% 
  # HACK! #############
  mutate(logor_hat = est_p_surv_no_HCQ, 
         logor_95_lower = est_p_surv_no_HCQ_95_lower, 
         logor_95_upper = est_p_surv_no_HCQ_95_upper) %>%
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = c(0, 1)) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous( 
    breaks = risk_breaks,
    labels = scales::percent_format(accuracy = 1)) + 
    ylab("survival rate under control") +
    ggtitle("Estimated survival rate under control: pooled and subgroup analyses"))

ggsave(filename = file.path(subgroup_dir, paste0("figure_p_surv_no_HCQ.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_p_surv_no_HCQ.svg")), width = 9, height = 10)
```

### Add mortality (1 - survival)

```{r fig.width=9, fig.height=10}
risk_breaks <- seq(0, 1, .2)

(fig_p_mortality_HCQ <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE)) %>% 
  # HACK! #############
  mutate(logor_hat = 1 - est_p_surv_HCQ, 
         logor_95_lower = 1 - est_p_surv_HCQ_95_upper, # NOTA BENE
         logor_95_upper = 1 - est_p_surv_HCQ_95_lower) %>%
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = c(0, 1)) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous( 
    breaks = risk_breaks,
    labels = scales::percent_format(accuracy = 1)) + 
    ylab("mortality rate under HC/CQ") +
    ggtitle("Estimated mortality rate under HC/CQ: pooled and subgroup analyses"))

ggsave(filename = file.path(subgroup_dir, paste0("figure_p_mortality_HCQ.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_p_mortality_HCQ.svg")), width = 9, height = 10)
```

```{r fig.width=9, fig.height=10}
risk_breaks <- seq(0, 1, .2)

(fig_p_mortality_no_HCQ <- all_effect_summaries  %>%
   mutate(subgroups = factor(subgroups, levels = grouping_order, ordered = TRUE),
          level = factor(level, levels = rev(level_order), ordered = TRUE)) %>% 
  # HACK! #############
  mutate(logor_hat = 1 - est_p_surv_no_HCQ, 
         logor_95_lower = 1 - est_p_surv_no_HCQ_95_upper, # NOTA BENE
         logor_95_upper = 1 - est_p_surv_no_HCQ_95_lower) %>%
   filter(analysis != "matched_posterior", level != "NA") %>%
   effect_summary_plot_sized_dots(
     all_group_sizes, 
     level) +
   coord_flip(ylim = c(0, 1)) +
   facet_grid(subgroups~., space = "free_y", scales = "free_y") +
   background_grid(major = "x", color.major = "white") +
   scale_color_manual(
     values =
       c("maximum likelihood" = "grey30",
         "superpop_posterior" = "cornflowerblue")
  ) + scale_y_continuous( 
    breaks = risk_breaks,
    labels = scales::percent_format(accuracy = 1)) + 
    ylab("mortality rate under control") +
    ggtitle("Estimated mortality rate under control: pooled and subgroup analyses"))

ggsave(filename = file.path(subgroup_dir, paste0("figure_p_mortality_no_HCQ.png")), width = 9, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_p_mortality_no_HCQ.svg")), width = 9, height = 10)
```

### Combined Figure 2

```{r}
require(patchwork)
```

```{r fig.height=10, fig.width=18}
figure_2_subgroups + fig_rd

ggsave(filename = file.path(subgroup_dir, paste0("figure_2_combined.png")), width = 18, height = 10)
ggsave(filename = file.path(subgroup_dir, paste0("figure_2_combined.svg")), width = 18, height = 10)
```







```{r}
sessionInfo()
```



```{r}
Sys.time()
```

