---
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: 
    html_document:
      keep_md: false
params:
  fit_name: "main_fit"
  outcome_min: 28
  outcome_max: 35
title: "`r paste0('Conditional effects - ', params$out_stub, params$fit_name, sep = '-')`"
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(here)
here::i_am(file.path("hcq_pooling_analysis", "4c_Conditional_effects.Rmd"))
source(here("hcq_pooling_analysis", "common.R"))
require(brms)
require(tidybayes)
require(mice)
require(bayesplot)
require(patchwork)

bayesplot_theme_set(theme_cowplot())
bayesplot_theme_update(
  strip.background = element_blank(),
  strip.text.y = element_text(angle = 0),
  strip.text = element_text(face = "bold")
)

require(MASS, exclude = "select") # clashes with dplyr

out_stub         <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir       <- here("hcq_pooling_analysis", "output", out_stub)
output_model_dir <- file.path(output_dir, params$fit_name)
fit_file         <- file.path(output_model_dir, paste0(params$fit_name, ".rds"))

conditional_effect_dir <-
  file.path(output_model_dir, "conditional_effects")

if (!dir.exists(conditional_effect_dir)) {
  dir.create(conditional_effect_dir)
}
```

```{r}
brm_fit <- read_rds(fit_file)
```

Create a reference table; this will be expanded for each conditional effect:

```{r}
reference_tbl <-
  tibble(
    treat = factor(c("no_HCQ", "HCQ"), levels = c("no_HCQ", "HCQ")),
    sex_model = 0,
    age_model = 0,
    bmi_model = 0,
    comorbidity_count = 0,
    niaid_baseline_numeric_model = 0,
    niaid_baseline_fct = "5 - hosp, no ox",
    
    # For expanded SAP model; otherwise unused
    azithro = FALSE,
    sym_onst_days_bfr_enrdt = 6 # the median
  )
```

### Generic functions for continuous covariates

```{r}
y_scale_percent <-
  scale_y_continuous(limits = c(0, 1), 
                     labels = scales::percent,
                     breaks = (1:4)/4,
                     expand = expansion(0))

conditional_analysis_tables <- 
  function(reference_table, model_fit, covariate_display) {
    conditional_tbl_pis <-
      add_fitted_draws(
        model_fit, newdata = reference_table ,
         # marginalize over contextual study effects
         re_formula = ~niaid_baseline_fct,
         scale = "response",
         # from the documentation for `add_fitted_draws` it seems
         # this might be necessary (though possibly only if we were adding
         # a dummy `studyid`)
          allow_new_levels = TRUE
         )
  
    conditional_tbl_12_rr <-
      conditional_tbl_pis %>%
      mutate(.category = as.numeric(substr(.category, 1, 2))) %>%
      group_by({{covariate_display}}, .draw) %>%
      summarise(prob_12_HCQ    = sum(.value*(treat == "HCQ")   *((.category == 1) + (.category == 2))),
                prob_12_no_HCQ = sum(.value*(treat == "no_HCQ")*((.category == 1) + (.category == 2))),
                rr = prob_12_HCQ/prob_12_no_HCQ)
    
    condition_tbl_linpred_diff <-
      add_linpred_draws(brm_fit, newdata = reference_table, 
                        re_formula = ~niaid_baseline_fct, scale = "linear") %>%
      ungroup() %>%
      group_by({{covariate_display}}, .draw) %>%
      summarise(lin_pred_diff = .value[treat == "HCQ"] - .value[treat == "no_HCQ"])
      # summarise(lin_pred_diff = .linpred[treat == "HCQ"] - .linpred[treat == "no_HCQ"])
    
    return(list(pi_tbl = conditional_tbl_pis,
                rr_tbl = conditional_tbl_12_rr,
                linpred_diff_tbl = condition_tbl_linpred_diff))
  }

pi_plot <- function(pi_draws, covariate_display, ref_level_display = NULL) {
  pi_draws %>%
  mutate(.category = as.numeric(substr(.category, 1, 2))) %>%
  filter(.category %in% c(1,2)) %>%
  group_by({{covariate_display}}, treat, .draw) %>%
  mutate(prob_12 = sum(.value)) %>%
  mutate(
    treat = factor(treat,
                   levels = c("no_HCQ", "HCQ"), 
                   labels = c("control", "HC/CQ"))) %>%
  ggplot(aes({{covariate_display}}, prob_12)) +
    facet_grid(treat~.) +
    stat_lineribbon(.width = c(.5, .95)) +
    scale_fill_brewer(palette = "Blues") +
    y_scale_percent +
    scale_x_continuous(expand = expansion(0)) +
    ylab("") +
    theme(panel.border = element_rect(color = "black", size = 1),
          axis.line = element_blank()) +
    geom_hline(yintercept = c(.25, .5, .75), color = "white") +
  geom_vline(xintercept = ref_level_display, 
             color = colorspace::lighten("forestgreen"), 
             size = 1) }

rr_plot <- function(rr_draws, covariate_display, ref_level_display = NULL) {
  rr_draws %>% 
  ggplot(aes({{covariate_display}}, rr)) + 
  stat_lineribbon(.width = c(.5, .95)) + 
  scale_y_log10() +
  scale_fill_brewer(palette = "Blues") + 
  geom_hline(yintercept = 1, color = "orange", size = 1) +
  ylab("relative risk of mechanical\nventilation or death\n(reference individuals)") +
  scale_x_continuous(expand = expansion(0)) +
  geom_vline(xintercept = ref_level_display, 
             color = colorspace::lighten("forestgreen"), 
             size = 1)
}

lin_pred_plot <-
  function(linpred_diff_draws, covariate_display, ref_level_display = NULL) {
    linpred_diff_draws %>%
    ggplot(aes({{covariate_display}}, lin_pred_diff)) + 
  stat_lineribbon(.width = c(.5, .95)) + 
  scale_fill_brewer(palette = "Blues") + 
  geom_hline(yintercept = 0, color = "orange", size = 1) +
  ylab("treatment effect (log proportional odds ratio) \n(reference individuals)") +
  scale_x_continuous(expand = expansion(0)) +
  geom_vline(xintercept = ref_level_display, 
             color = colorspace::lighten("forestgreen"), 
             size = 1)
  }
```

### BMI

```{r}
conditional_tables_bmi <-
  tidyr::expand_grid(
      reference_tbl,
      bmi = seq(15, 80, 5)
    ) %>%
    mutate(bmi_model = (bmi - 25)/5) %>%
  conditional_analysis_tables(brm_fit, bmi)
```



```{r fig.height=12, fig.width=6}
bmi_plots_stacked <-
  ((rr_plot(conditional_tables_bmi$rr_tbl, bmi, ref_level_display = 25) + scale_y_log10()) /
  pi_plot(conditional_tables_bmi$pi_tbl, bmi, ref_level_display = 25) /
  lin_pred_plot(conditional_tables_bmi$linpred_diff_tbl, bmi, ref_level_display = 25)) +
  plot_annotation("Conditional effects - BMI")

bmi_plots_stacked
ggsave(file.path(conditional_effect_dir, "bmi_stacked.png"), bmi_plots_stacked, height = 12, width = 6)
ggsave(file.path(conditional_effect_dir, "bmi_stacked.svg"), bmi_plots_stacked, height = 12, width = 6)
```


### Baseline NIAID

```{r}
conditional_tables_baseline_niaid <-
  reference_tbl %>%
  select(-niaid_baseline_fct, -niaid_baseline_numeric_model) %>%
  tidyr::expand_grid(
      tibble(
        niaid_baseline_fct = niaid_levels[2:5],
        niaid_baseline_numeric_model = 3:0
      )
    ) %>%
  conditional_analysis_tables(brm_fit, niaid_baseline_fct)
```

### Functions for discrete covariates

```{r}
pi_plot_discrete <- function(pi_draws, covariate_display, ref_level_display) {
  x_val <- ref_level_display
  alpha_val <- .01

  if(!is.null(ref_level_display)) {
  pi_draws %>%
    mutate(.category = as.numeric(substr(.category, 1, 2))) %>%
    filter(.category %in% c(1,2)) %>%
    group_by({{covariate_display}}, treat, .draw) %>%
    mutate(prob_12 = sum(.value)) %>%
    mutate(
      treat = factor(treat,
                     levels = c("no_HCQ", "HCQ"), 
                     labels = c("control", "HC/CQ"))) %>%
    ggplot(aes({{covariate_display}}, prob_12)) +
      background_grid(major = "y", color.major = "grey90") +
      facet_rep_grid(treat~.) +
    geom_tile(
      aes(x = x_val, 
          y = .5, height = Inf, width = 1), 
      alpha = alpha_val, 
      fill = colorspace::lighten("forestgreen")) +
    stat_pointinterval(.width = c(.5, .95)) +
    # scale_fill_brewer(palette = "Greys") +
    y_scale_percent +
    # scale_x_continuous(expand = expansion(0)) +
    ylab("probability of mechanical\nventilation or death\n(reference individuals)")
    # theme(panel.border = element_rect(color = "black", size = 1),
    #       axis.line = element_blank()) +
    # geom_hline(yintercept = c(.25, .5, .75), color = "white") 
  } else {
    pi_draws %>%
    mutate(.category = as.numeric(substr(.category, 1, 2))) %>%
    filter(.category %in% c(1,2)) %>%
    group_by({{covariate_display}}, treat, .draw) %>%
    mutate(prob_12 = sum(.value)) %>%
    ggplot(aes({{covariate_display}}, prob_12)) +
      background_grid(major = "y", color.major = "grey90") +
      facet_rep_grid(treat~.) +
    # geom_tile(
    #   aes(x = x_val, 
    #       y = .5, height = Inf, width = 1), 
    #   alpha = alpha_val, 
    #   fill = colorspace::lighten("forestgreen")) +
    stat_pointinterval(.width = c(.5, .95)) +
    # scale_fill_brewer(palette = "Greys") +
    y_scale_percent +
    # scale_x_continuous(expand = expansion(0)) +
    ylab("probability of mechanical\nventilation or death\n(reference individuals)")
    # theme(panel.border = element_rect(color = "black", size = 1),
    #       axis.line = element_blank()) +
    # geom_hline(yintercept = c(.25, .5, .75), color = "white") 
  }
}

pi_plot_discrete(conditional_tables_baseline_niaid$pi_tbl, niaid_baseline_fct, 
                 ref_level_display = "5 - hosp, no ox"
                 )
```

```{r}
rr_plot_discrete <- function(rr_draws, covariate_display, ref_level_display) {
  x_val <- ref_level_display
  alpha_val <- .01
  
  if (!is.null(ref_level_display)) {
  rr_draws %>% 
  ggplot(aes({{covariate_display}}, rr)) +
  geom_tile(
    aes(x = x_val, 
        y = .5, height = Inf, width = 1), 
    alpha = alpha_val, 
    fill = colorspace::lighten("forestgreen")) + 
  stat_pointinterval(.width = c(.5, .95)) +
  scale_y_log10(breaks = scales::log_breaks()) +
  # scale_fill_brewer(palette = "Greys") + 
  geom_hline(yintercept = 1, color = "orange", size = .5) +
  ylab("relative risk of mechanical\nventilation or death\n(reference individuals)")
  # scale_x_continuous(expand = expansion(0))
  } else {
  rr_draws %>% 
  ggplot(aes({{covariate_display}}, rr)) +
  stat_pointinterval(.width = c(.5, .95)) +
  scale_y_log10(breaks = scales::log_breaks()) +
  # scale_fill_brewer(palette = "Greys") + 
  geom_hline(yintercept = 1, color = "orange", size = .5) +
  ylab("relative risk of mechanical\nventilation or death\n(reference individuals)")
  }
}

rr_plot_discrete(conditional_tables_baseline_niaid$rr_tbl, niaid_baseline_fct, ref_level_display = NULL)
```

```{r}
linpred_plot_discrete <- function(linpred_diff_draws, covariate_display, ref_level_display) {
  x_val <- ref_level_display
  alpha_val <- .01
  
  if(!is.null(ref_level_display)) {
  linpred_diff_draws %>% 
  ggplot(aes({{covariate_display}}, lin_pred_diff)) +
  geom_tile(
    aes(x = x_val, 
        y = 0, height = Inf, width = 1), 
    alpha = alpha_val, 
    fill = colorspace::lighten("forestgreen")) + 
  stat_pointinterval(.width = c(.5, .95)) +
  # scale_fill_brewer(palette = "Greys") + 
  geom_hline(yintercept = 0, color = "orange", size = .5) +
  ylab("treatment effect\n(log proportional odds ratio) \n(reference individuals)")
  # scale_x_continuous(expand = expansion(0))
  } else {
    linpred_diff_draws %>% 
  ggplot(aes({{covariate_display}}, lin_pred_diff)) +
  stat_pointinterval(.width = c(.5, .95)) +
  # scale_fill_brewer(palette = "Greys") + 
  geom_hline(yintercept = 0, color = "orange", size = .5) +
  ylab("treatment effect\n(log proportional odds ratio) \n(reference individuals)")
  # scale_x_continuous(expand = expansion(0))
  }
}

linpred_plot_discrete(conditional_tables_baseline_niaid$linpred_diff_tbl, niaid_baseline_fct,
                      "5 - hosp, no ox")
```

```{r fig.height=12, fig.width=6}
x <- theme(axis.text.x = element_text(angle = 90, vjust = .5))
y <- theme(axis.text.x = element_blank())

niaid_plot_stacked <-
  ((rr_plot_discrete(conditional_tables_baseline_niaid$rr_tbl, niaid_baseline_fct, ref_level_display = "5 - hosp, no ox") + y) /
  (pi_plot_discrete(conditional_tables_baseline_niaid$pi_tbl, niaid_baseline_fct, 
                 ref_level_display = "5 - hosp, no ox"
                 ) + ylab("probability of mechanical\nventilation or death\n(reference individuals)") + y) /
  (linpred_plot_discrete(conditional_tables_baseline_niaid$linpred_diff_tbl, niaid_baseline_fct,
                      "5 - hosp, no ox") + x)) +
  plot_annotation("Conditional effects - Baseline NCOSS")

niaid_plot_stacked

ggsave(file.path(conditional_effect_dir, "niaid_stacked.png"), niaid_plot_stacked, height = 12, width = 6)
ggsave(file.path(conditional_effect_dir, "niaid_stacked.svg"), niaid_plot_stacked, height = 12, width = 6)
```

Finish compiling conditional effects:

-   age (done)
-   BMI (done)
-   baseline NCOSS (done)



### Sex

We don't have a reference level here.

```{r}
conditional_tables_sex <-
  tidyr::expand_grid(
      reference_tbl %>% select(-sex_model),
      sex_fct = c("male", "female")
    ) %>%
  mutate(sex_model = ifelse(sex_fct == "male", -.5, .5)) %>%
  conditional_analysis_tables(brm_fit, sex_fct)
  
```

```{r fig.height=12, fig.width=6}
sex_plots_stacked <-
  (rr_plot_discrete(conditional_tables_sex$rr_tbl, sex_fct, ref_level_display = NULL) /
  (pi_plot_discrete(conditional_tables_sex$pi_tbl, sex_fct, ref_level_display = NULL
                 ) + ylab("probability of mechanical\nventilation or death\n(reference individuals)")) /
  linpred_plot_discrete(conditional_tables_sex$linpred_diff_tbl, sex_fct, ref_level_display = NULL)) +
  plot_annotation(title = "Conditional effects - sex")

sex_plots_stacked
ggsave(file.path(conditional_effect_dir, "sex_stacked.png"), sex_plots_stacked, height = 12, width = 6)
ggsave(file.path(conditional_effect_dir, "sex_stacked.svg"), sex_plots_stacked, height = 12, width = 6)
```

### Number of baseline comorbidities

Treat this like a continuous covariate:

```{r}
conditional_tables_comorbidities <-
  tidyr::expand_grid(
      reference_tbl %>% select(-comorbidity_count),
      comorbidity_count = seq(0, 9, .5)
    ) %>%
  conditional_analysis_tables(brm_fit, comorbidity_count)
```

```{r fig.height=12, fig.width=6}
comorbidity_plots_stacked <-
  ((rr_plot(conditional_tables_comorbidities$rr_tbl, comorbidity_count, ref_level_display = 0) + scale_y_log10() + scale_x_continuous(breaks = 0:9)) /
  (pi_plot(conditional_tables_comorbidities$pi_tbl, comorbidity_count, ref_level_display = 0)
    + scale_x_continuous(breaks = 0:9)) /
  (lin_pred_plot(conditional_tables_comorbidities$linpred_diff_tbl, comorbidity_count, ref_level_display = 0)) + scale_x_continuous(breaks = 0:9)) +
  plot_annotation("Conditional effects - Baseline comorbidity count")

comorbidity_plots_stacked
ggsave(file.path(conditional_effect_dir, "comorbidity_stacked.png"), comorbidity_plots_stacked, height = 12, width = 6)
ggsave(file.path(conditional_effect_dir, "comorbidity_stacked.svg"), comorbidity_plots_stacked, height = 12, width = 6)
```

### Age

```{r}
conditional_tables_age <-
  tidyr::expand_grid(
      reference_tbl,
      age = seq(18, 90, 3)
    ) %>%
    mutate(age_model = (age - 60)/10) %>%
  conditional_analysis_tables(brm_fit, age)
```

```{r fig.height=12, fig.width=6}
age_plots_stacked <-
  ((rr_plot(conditional_tables_age$rr_tbl, age, ref_level_display = 60) + 
      scale_y_log10()) /
  (pi_plot(conditional_tables_age$pi_tbl, age, ref_level_display = 60)) /
  (lin_pred_plot(conditional_tables_age$linpred_diff_tbl, age, ref_level_display = 60))) +
  plot_annotation("Conditional effects - Age")

age_plots_stacked
ggsave(file.path(conditional_effect_dir, "age_stacked.png"), age_plots_stacked, height = 12, width = 6)
ggsave(file.path(conditional_effect_dir, "age_stacked.svg"), age_plots_stacked, height = 12, width = 6)
```





```{r}
sessionInfo()
```



```{r}
Sys.time()
```

