---
title: "Table 1 for paper"
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: html_document
params:
  fit_name: "main_fit"
  outcome_min: 28
  outcome_max: 35
---

## Table 1 for paper

```{r}
knitr::opts_chunk$set(echo = TRUE)
require(here)
here::i_am(file.path("hcq_pooling_analysis", "6_Table_1.Rmd"))
source(here("hcq_pooling_analysis", "common.R"))

require(table1)
out_stub         <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir       <- here("hcq_pooling_analysis", "output", out_stub)
output_model_dir <- file.path(output_dir, params$fit_name)
```

Need to

-   Compute comorbidity sum (used in the model)
-   Choose the correct azithro variables (also corticosteroids)
-   Figure out exactly which variables to include
-   Format their names nicely
-   ✓ Simplify the continuous statistics to "median (q1 to q3)" + missingness

Function to use medians and IQRs:

```{r}
render_continuous_custom <- function (x, ...) 
{
    with(stats.apply.rounding(
      stats.default(x), 
      digits = 1, 
      digits.pct = 0, 
      rounding.fn = round_pad # Decimal places rather than significant digits
      ), c("", 
        `mean (SD)` = sprintf("%s (%s)", MEAN, SD), 
        `median (IQR)` = sprintf("%s (%s to %s)", 
            MEDIAN, Q1, Q3)))
}

render_categorical_custom <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%0.f)", FREQ, PCT))))
}
render_missing_custom <- function (x, ...) 
{
    with(stats.apply.rounding(stats.default(is.na(x), ...), digits.pct = 0, rounding.fn = round_pad)$Yes, 
        c(missing = sprintf("%s (%s%%)", FREQ, PCT)))
}
```

Need to load `data_tbl` to get the comorbidity count variable used in the model:

```{r}
baseline_formatted <- read_rds(file.path(output_dir, "patients.rds"))
data_tbl           <- read_rds(file.path(output_dir, "data_tbl.rds"))
```

```{r}
nrow(baseline_formatted); nrow(data_tbl)
ncol(baseline_formatted); ncol(data_tbl)
```

```{r}
combined_tbl <- 
  left_join(
    data_tbl %>% select(comorbidity_count, patient_id, niaid_outcome),
    baseline_formatted %>% select(-niaid_outcome),
    by = "patient_id")

dim(combined_tbl)
```

Create azithro variable:

```{r}
combined_tbl <-
  combined_tbl %>% 
  mutate(
     azithro_use = case_when(
       azithro    & on_azithromycin    ~ "assigned, took",
       (!azithro) & on_azithromycin    ~ "not assigned, took (on or before d28)",
       (!azithro) & (!on_azithromycin) ~ "not assigned, did not take",
       azithro    & (!on_azithromycin) ~ "assigned, did not take (on or before d28)",
     ),
     covid_scale_baseline_fct = niaid_baseline_fct,
     covid_scale_baseline_numeric = as.numeric(niaid_baseline), # original Stata encoding is correct
     on_corticosteroids = as.logical(on_corticosteroids),
     first_dose_at_enrollment = frstdose_days_after_enrdt == 0,
     outcome_missing = is.na(niaid_outcome))
```

Set labels:

```{r}
label(combined_tbl$sex_fct) <- "Sex"
label(combined_tbl$race_simplified_fct) <- "Race (simplified)"
label(combined_tbl$ethnic_fct) <- "Ethnicity"
label(combined_tbl$age_5y) <- "Age (5 year bins)"
label(combined_tbl$bmi) <- "BMI"
label(combined_tbl$covid_scale_baseline_fct) <- "Baseline NCOSS"
label(combined_tbl$covid_scale_baseline_numeric) <- "Baseline NCOSS (numeric)"
label(combined_tbl$sym_onst_days_bfr_enrdt) <- "Days between symptom onset and enrollment"
label(combined_tbl$siteid) <- "Site"
label(combined_tbl$treat) <- "Treatment group"

label(combined_tbl$comorbidity_count) <- "Baseline comorbidity count"

label(combined_tbl$azithro_use) <- "Azithromycin use"
label(combined_tbl$on_corticosteroids) <- "Concurrent corticosteroid use (on or before d28)"
label(combined_tbl$first_dose_at_enrollment) <- "First dose recieved on day of enrollment"

label(combined_tbl$outcome_missing) <- "Missing outcome (NCOSS between d28-d35)"
```


```{r}
table_1 <-
  table1(~ sex_fct + 
         race_simplified_fct + ethnic_fct + 
         age_5y +
         bmi +
         covid_scale_baseline_fct +
         covid_scale_baseline_numeric +
         sym_onst_days_bfr_enrdt +
           
         comorbidity_count +
         
         azithro_use +
         on_corticosteroids +
           
         first_dose_at_enrollment +
         
         outcome_missing
       | siteid * treat, 
       data = combined_tbl,
       render.continuous  = render_continuous_custom,
       render.categorical = render_categorical_custom,
       render.missing     = render_missing_custom,
       topclass = "Rtable1-grid Rtable1-center")

table_1
```

```{r}
require(kableExtra)
table1::t1kable(table_1)
```










```{r}
sessionInfo()
```



```{r}
Sys.time()
```

