---
title: "8 Misc manuscript numbers"
author: "Leon Di Stefano"
date: "8/20/2021"
output: html_document
params:
  fit_name: "main_fit"
  outcome_min: 28
  outcome_max: 35
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(here)
here::i_am(file.path("hcq_pooling_analysis", "7_Misc_manuscript_numbers.Rmd"))

require(meta)
```

```{r}
out_stub         <- paste(params$outcome_min, params$outcome_max, sep = '-')
output_dir       <- here("hcq_pooling_analysis", "output", out_stub)
output_model_dir <- file.path(output_dir, params$fit_name)

data_tbl <- read_rds(file.path(output_dir, "data_tbl.rds"))
patients <- read_rds(file.path(output_dir, "patients.rds")) %>% filter(passed_screen == 1)
study_arms <- read_tsv("resources/study_arm_tbl.csv")
```

Number of patients:

```{r}
nrow(data_tbl)
```

Treat vs control

```{r}
count(data_tbl, treat)
```

Mortality:

```{r}
data_tbl %>% count(treat, as.numeric(niaid_outcome) == 1) %>% group_by(treat) %>% mutate(pct = 100 * n/sum(n))
```

Median onset days before enrdt:

```{r}
summary(data_tbl$sym_onst_days_bfr_enrdt)
```

```{r}
hist(data_tbl$sym_onst_days_bfr_enrdt, 40)
```

Dose information: median (IQR) length of dosing among the HCQ patients (controls not comparable):

```{r}
patients %>% 
    mutate(
      dose_duration = 
        lstdose_days_after_enrdt - frstdose_days_after_enrdt) %>% 
  group_by(treat) %>% 
  summarise(
    val = quantile(
      dose_duration, 
      c(.25, .5, .75), 
      na.rm = TRUE), 
    quantile = names(val))
```

Percent beginning dosing on day 0:

```{r}
patients %>%
  bind_rows(patients %>% mutate(treat = "OVERALL")) %>%
  group_by(treat) %>%
  summarise(mean(frstdose_days_after_enrdt == 0, na.rm = TRUE),
            sum(frstdose_days_after_enrdt == 0, na.rm = TRUE),
            sum(is.na(frstdose_days_after_enrdt)))
```

```{r}
sum(is.na(patients$frstdose_days_after_enrdt))
```

HCQ vs. CQ:

```{r}
study_arms
```

```{r}
count(patients, treat)
```

```{r}
count(patients, treat, CQ)
```

Corticosteroids:


Missingness at for different outcome definitions (taking into account mortality inconsistencies not known at the design stage):

```{r}
x35 <- read_rds("output/28-35/data_tbl.rds")
x30 <- read_rds("output/28-30/data_tbl.rds")

round(100 * mean(!is.na(x35$niaid_outcome)), 2)
round(100 * mean(!is.na(x30$niaid_outcome)), 2)
```

Comparing ORCHID-only versus full fit CI widths:

```{r}
orchid_only_subgroups <- read_tsv("output/28-35/main_fit_Vanderbilt_only/subgroup_effects/subgroup_analysis_table.tsv")
full_model_subgroups  <- read_tsv("output/28-35/main_fit/subgroup_effects/subgroup_analysis_table.tsv")

orchid_vs_full_CI_widths <- 
  bind_cols(
    full_model_subgroups %>% select(analysis, subgroups, level),
    tibble(
  orchid_CI_width = orchid_only_subgroups %>% transmute(or_CI_width = logor_95_upper - logor_95_lower) %>% pull(or_CI_width),
  full_CI_width   = full_model_subgroups  %>% transmute(or_CI_width = logor_95_upper - logor_95_lower) %>% pull(or_CI_width)
)) %>% filter(analysis == "superpop_posterior")

orchid_vs_full_CI_widths
```

```{r}
orchid_vs_full_CI_widths %>%
  qplot(orchid_CI_width, full_CI_width, data = .) +
  geom_abline()
```

```{r}
orchid_vs_full_CI_widths %>%
  filter(orchid_CI_width < full_CI_width)
```

Adjusted OR from additive model:

```{r}
additive_fit <- read_rds("output/28-35/main_fit_no_interactions/main_fit_no_interactions.rds")
```

```{r}
additive_summary <- summary(additive_fit)

additive_summary$fixed["treatHCQ",c("Estimate", "l-95% CI", "u-95% CI")] %>%
  exp() %>%
  round(2)
```

Pre-analysis tweaks of values (BMI and possibly—though not in fact—symptoms onset days before enrollment):

```{r}
patients %>% count(sym_onst_days_bfr_enrdt < 0)
```

```{r}
patients %>% count(bmi < 10 | bmi > 70)
```

```{r}
data_tbl %>% count(bmi < 10 | bmi > 70)
```

# Begg and Egger tests

From plugin subgroup analysis by site:

```{r}
subgroup_analysis_tbl <- 
  read_tsv(
    file.path(
      output_model_dir,
      "subgroup_effects/subgroup_analysis_table.tsv"))

(site_plugin_tbl <-
  subgroup_analysis_tbl %>%
  filter(subgroups == "site", analysis == "maximum likelihood") %>%
  select(level, logor_hat, logor_se))
```

```{r}
(site_plugin_meta_object <- 
  meta::metagen(
    TE = site_plugin_tbl$logor_hat, 
    seTE = site_plugin_tbl$logor_se))
```

Begg test:

```{r}
metabias(
  site_plugin_meta_object,
  plotit = TRUE,
  k.min = 6,
  method.bias = "rank" # Begg
)
```

Egger test:

```{r}
metabias(
  site_plugin_meta_object,
  plotit = TRUE,
  k.min = 6,
  method.bias = "linreg" # Egger
)
```

### Exploratory analysis of days from symptom onset to enrollment

```{r}
require(patchwork)
require(cowplot)
theme_set(theme_cowplot())

sym_onset_subgroup_tbl <- 
  read_tsv("output/28-35/main_fit/subgroup_effects/subgroup_analysis_table.tsv") %>%
  filter(
    subgroups == "sym_onset_fct",
    analysis == "maximum likelihood",
    !is.na(level)) %>%
  mutate(
    level = 
      factor(level, 
             labels = str_c((level), "\n(n=", n, ")")) %>%
      fct_rev())
```

ORs:

```{r}
(or_plot <- sym_onset_subgroup_tbl %>%
  ggplot(aes(level, exp(logor_hat))) +
  geom_hline(yintercept = 1, lty = "dotted") +
  geom_pointrange(aes(ymin = exp(logor_95_lower), ymax = exp(logor_95_upper))) +
  ylab("Proportional odds ratio") +
  scale_y_log10(labels = function(s) str_c(s, "×")))
```

RDs:

```{r}
(surv_rd_plot <- sym_onset_subgroup_tbl %>%
  ggplot(aes(level, rd_surv)) +
  geom_hline(yintercept = 0, lty = "dotted") +
  geom_pointrange(aes(ymin = rd_surv_95_lower, ymax = rd_surv_95_upper)) +
  ylab("Risk difference for mortality") +
  scale_y_continuous(labels = function(x) str_c(100*x, "%")))
```



```{r}
(combined_risk_plot <-
  sym_onset_subgroup_tbl %>%
  ggplot(aes(level)) +
  geom_hline(yintercept = 0, lty = "dotted") +
  geom_pointrange(aes(
    y = 1 - est_p_surv_no_HCQ, 
    ymin = 1 - est_p_surv_no_HCQ_95_upper, # NOTA BENE 
    ymax = 1 - est_p_surv_no_HCQ_95_lower),
    position = position_nudge(x = -.12),
    color = "grey40",
    fill = "white",
    shape = 22) + # Square
  geom_pointrange(aes(
    y = 1 - est_p_surv_HCQ, 
    ymin = 1 - est_p_surv_HCQ_95_upper, 
    ymax = 1 - est_p_surv_HCQ_95_lower),
    position = position_nudge(x = +.12),
    color = "grey40",
    fill = "grey40",
    shape = 22) + # Triangle
  ylab("Mortality") +
  scale_y_continuous(
    labels = (function(x) str_c(100*x, "%")),
    limits = c(0, NA),
    expand = expansion(c(0, .5))
   )
)
```

```{r fig.width=10, fig.height=3.5}
theme_update(axis.title.y = element_text(angle = 0, vjust = .5))
(combined_plot <-
  (or_plot + 
   coord_flip() +
   xlab("Days between\nsymptom onset\nand enrollment")
 ) + 
  (combined_risk_plot +
     coord_flip() + 
     theme(axis.text.y = element_blank(), axis.title.y = element_blank()) +
     cowplot::background_grid("x", "none", colour.major = "grey90")
   ) +
  (surv_rd_plot + 
     coord_flip() + 
     theme(axis.text.y = element_blank(), axis.title.y = element_blank())
   ) +
  patchwork::plot_layout(nrow = 1))

ggsave("output/28-35/sym_onset_sensitivity_analysis.svg")
```

```{r}
sessionInfo()
```

```{r}
Sys.time()
```

