---
title: "Final design-stage checks"
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output: 
    html_document:
      self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(tidyverse)
require(cowplot)
require(lemon)
theme_set(theme_cowplot())
theme_update(
  strip.background = element_blank(),
  strip.text.y = element_text(angle = 0),
  )

require(ggbeeswarm)
gg_scatter <- function(data, x_var, y_var) {
  qplot({{x_var}}, {{y_var}}, 
        data = data,
        size = I(0.4)) +
    cowplot::theme_cowplot()
}

gg_hist <- function(data, x_var, n_bins = 30) {
  qplot({{x_var}}, bins = I(n_bins), 
        data = data,
        color = I("white")) +
    scale_y_continuous(expand = expansion(0, .05)) +
    cowplot::theme_cowplot()
}

require(naniar)
require(table1)
require(gt)
require(haven)
```

## Load the data

```{r}
patient_raw <- read_dta(
  "../data/2021-09-29_HCQ_Analysis_Datasets/HCQ_Analysis_Datasets/8Site_PatientData.dta")

baseline_raw <- read_dta(
  "../data/2021-09-29_HCQ_Analysis_Datasets/HCQ_Analysis_Datasets/8Sites_Outcome_BL.dta")

# Check that all of the baseline values have "days post enrollment" == 0
all(baseline_raw$days_post_enrollment == 0)
```

Need to deal with non-unique patient ids between sites:

```{r}
baseline_raw$patient_id %>% unique() %>% length()
nrow(baseline_raw)
baseline_raw %>% select(siteid, patient_id) %>% distinct() %>% nrow()
```

```{r}
combined_all <- 
  left_join(
    patient_raw %>%
      filter(passed_screen == 1) %>%
      mutate(patient_id = str_c(siteid, patient_id, sep = "__")), 
    baseline_raw %>%
      mutate(# FIXED
             # siteid = str_replace(siteid, "3-DUke", "3-Duke"),
             patient_id = str_c(siteid, patient_id, sep = "__"))) %>%
  select(-days_post_enrollment) %>%
  rename(covid_scale_baseline = covid_scale)

nrow(combined_all)
```

These are the final design-stage checks as described in our SAP.

## Summary of missingness

> We will also examine baseline covariates for missing data (See section 8).


> We will assess the extent of missing baseline covariate data at the design stage. If the missingness
is judged to be minor and sporadic, the missing values will be filled-in using multiple imputation.
If the missingness is judged to be severe, the corresponding baseline variables will be excluded
from the analysis.

### Outcome missingness

```{r}
outcomes_uncensored <-
  read_dta(
    "../data/2021-09-29_HCQ_Analysis_Datasets/HCQ_Analysis_Datasets/8Sites_Outcome_Data.dta") %>%
  mutate(
         # FIXED:
         # siteid = 
         #   # Fix typos
         #   str_replace(siteid, "3-DUke", "3-Duke") %>%
         #   str_replace("^NYU$", "4-NYU-TEACH"),
         patient_id = 
           # FIXED
           # Typos in patient IDs
           # case_when(
           #   (siteid == "4-NYU-TEACH") & (str_length(patient_id) == 2) ~ str_c("0", patient_id),
           #   (siteid == "4-NYU-TEACH") & (str_length(patient_id) == 1) ~ str_c("00", patient_id),
           #   (siteid == "8-QMC-Honolulu") & (patient_id == "POB11")    ~ "PB11",
           #   TRUE                                                      ~ patient_id
           # ) %>%
           patient_id %>%
           str_c(siteid, ., sep = "__"))

# Any patient_ids IN combined all that are NOT in outcomes_unc
combined_all %>% anti_join(outcomes_uncensored) %>% nrow()

outcomes_uncensored <- filter(outcomes_uncensored, patient_id %in% combined_all$patient_id)

patient_time_grid <-
  expand_grid(
    patient_id = unique(outcomes_uncensored$patient_id),
    days_post_enrollment = 
      min(outcomes_uncensored$days_post_enrollment):max(outcomes_uncensored$days_post_enrollment)
  ) %>%
  arrange(patient_id, days_post_enrollment) %>%
  mutate(days_post_enrollment = as.numeric(days_post_enrollment)) %>%
  left_join(outcomes_uncensored) %>%
  group_by(patient_id) %>%
  mutate(
    died_before = cumsum(replace_na((covid_scale == 1), 0)) > 0,
    has_outcome_value =
      (!is.na(covid_scale)) | (died_before)) %>%
  select(patient_id, days_post_enrollment, has_outcome_value) %>%
  ungroup()

rm(outcomes_uncensored)

limits_tbl <-
  cross_df(
    list(start_day = 20:28, end_day = 30:40)
  )

n_patients_w_gt_one_measurement <- function(start_day, end_day) {
  filter(patient_time_grid,
         (days_post_enrollment >= start_day) & (days_post_enrollment <= end_day)) %>%
    group_by(patient_id) %>%
    summarise(any_measurements = any(has_outcome_value)) %>%
    filter(any_measurements == 1) %>%
    nrow()
}

(cases_by_range <-
    limits_tbl %>%
    mutate(n_patients_w_any_measurement = map2_dbl(start_day, end_day, n_patients_w_gt_one_measurement))) %>%
  pivot_wider(names_from = start_day, values_from = n_patients_w_any_measurement) %>%
  gt::gt(rowname_col = "end_day") %>%
  gt::tab_spanner("cols: first day; rows: last day", columns = as.character(20:28)) %>%
  gt::tab_header("# patients w/ outcome data")

```

```{r}
n_patients <- length(unique(patient_time_grid$patient_id))

cases_by_range %>%
  mutate(prop_patients_w_any_measurement = n_patients_w_any_measurement/n_patients,
         pct_patients_w_any_measurement = round(prop_patients_w_any_measurement * 100, 1)) %>%
  select(start_day, end_day, pct_patients_w_any_measurement) %>%
  pivot_wider(names_from = start_day, values_from = pct_patients_w_any_measurement) %>%
  gt::gt(rowname_col = "end_day") %>%
  gt::tab_spanner("cols: first day; rows: last day", columns = as.character(20:28)) %>%
  gt::tab_header("% patients w/ outcome data")
```

```{r}
cases_by_range %>%
  mutate(prop_patients_w_any_measurement = n_patients_w_any_measurement/n_patients,
         pct_patients_w_any_measurement = round(prop_patients_w_any_measurement * 100, 1)) %>%
  select(start_day, end_day, pct_patients_w_any_measurement) %>%
  pivot_wider(names_from = start_day, values_from = pct_patients_w_any_measurement) %>%
  gt::gt(rowname_col = "end_day") %>%
  gt::tab_spanner("cols: first day; rows: last day", columns = as.character(20:28)) %>%
  gt::tab_header("% patients w/ outcome data")
```

An expanded version of the above:

```{r, message=FALSE}
expanded_tbl <-
  cross_df(
    list(start_day = 0:40, end_day = 0:40)
  ) %>%
    mutate(n_patients_w_any_measurement = map2_dbl(start_day, end_day, n_patients_w_gt_one_measurement)) %>%
  mutate(prop_patients_w_any_measurement = n_patients_w_any_measurement/n_patients,
         pct_patients_w_any_measurement = round(prop_patients_w_any_measurement * 100, 1)) %>%
  select(start_day, end_day, pct_patients_w_any_measurement) %>%
  mutate(start_day = str_c("d", start_day)) %>%
  pivot_wider(names_from = start_day, values_from = pct_patients_w_any_measurement)

expanded_tbl %>%
  gt::gt(rowname_col = "end_day") %>%
  gt::tab_spanner("cols: first day; rows: last day", columns = str_c("d", as.character(20:28))) %>%
  gt::tab_header("% patients w/ outcome data") %>%
  gt::data_color(
    columns = str_c("d", as.character(0:40)),
    colors = scales::col_numeric(
      palette = c("lightgrey", "orange"),
      domain  = c(0, 100)
  ))
```

```{r}
cases_by_range %>%
  mutate(n_patients_wo_any_measurement = n_patients - n_patients_w_any_measurement) %>%
  select(start_day, end_day, n_patients_wo_any_measurement) %>%
  pivot_wider(names_from = start_day, values_from = n_patients_wo_any_measurement) %>%
  gt::gt(rowname_col = "end_day") %>%
  gt::tab_spanner("cols: first day; rows: last day", columns = as.character(20:28)) %>%
  gt::tab_header("# patients missing outcome data")
```

```{r fig.height=3, fig.width=4.5}
d28_dist_tbl <-
  patient_time_grid %>% 
  # add site information
  left_join(combined_all %>% select(patient_id, siteid)) %>%
  group_by(siteid, patient_id) %>% 
  mutate(
    d28_dist = ifelse(
      has_outcome_value, 
      abs(days_post_enrollment - 28),
      Inf)) %>% 
  summarise(d28_nearest_dist = min(d28_dist))

gg_hist(d28_dist_tbl, d28_nearest_dist)
```

```{r fig.height=8, fig.width=5}
gg_hist(d28_dist_tbl, d28_nearest_dist) +
  facet_rep_grid(siteid~., scales = "free_y") +
  scale_y_continuous(breaks = scales::pretty_breaks(2),
                     expand = expansion(c(0, .1))) +
  theme(
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0),
  )
```

```{r}
tibble(
  dist_from_d28 = 0:30,
  percent_w_measurement = 
    round(100 * ecdf(d28_dist_tbl$d28_nearest_dist)(dist_from_d28), 1),
  n_w_measurement = n_patients * ecdf(d28_dist_tbl$d28_nearest_dist)(dist_from_d28)
) %>% gt::gt()
```

```{r fig.width=19, fig.height=6}
patient_time_grid %>% 
  left_join(
    combined_all %>% select(patient_id, siteid)) %>% 
  left_join(d28_dist_tbl) %>%
  mutate(patient_id = factor(patient_id, levels = arrange(d28_dist_tbl, d28_nearest_dist)$patient_id, ordered = TRUE)) %>%
  ggplot(aes(patient_id, days_post_enrollment)) + 
  geom_tile(
    aes(fill = has_outcome_value), color = "white") + 
  facet_grid(.~siteid, scales = "free_x", space = "free_x") + 
  scale_fill_manual(values = c("grey90", "orange")) + 
  theme(
    axis.ticks.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.line.x = element_blank()) + 
  scale_y_continuous() +
  geom_hline(yintercept = 28 - .5, color = "red") +
  geom_hline(yintercept = 35 + .5, color = "red")
```

```{r fig.width=19, fig.height=6}
patient_time_grid %>% 
  left_join(
    combined_all %>% select(patient_id, siteid)) %>% 
  left_join(d28_dist_tbl) %>%
  mutate(patient_id = factor(patient_id, levels = arrange(d28_dist_tbl, d28_nearest_dist)$patient_id, ordered = TRUE)) %>%
  filter(siteid != "1-ORCHID") %>%
  ggplot(aes(patient_id, days_post_enrollment)) + 
  geom_tile(
    aes(fill = has_outcome_value), color = "white") + 
  facet_grid(.~siteid, scales = "free_x", space = "free_x") + 
  scale_fill_manual(values = c("grey90", "orange")) + 
  theme(
    axis.ticks.x = element_blank(), 
    axis.text.x = element_blank(), 
    axis.line.x = element_blank()) + 
  scale_y_continuous() +
  geom_hline(yintercept = 28 - .5, color = "red") +
  geom_hline(yintercept = 35 + .5, color = "red") +
  geom_hline(yintercept = 30 + .5, color = "red")
```

Table of missingness for each site and overall under the d28-30 and d28-35 outcome definitions:

```{r}
patient_time_grid %>% 
    left_join(
        combined_all %>% select(patient_id, siteid)) %>% 
  bind_rows(
    patient_time_grid %>% 
    left_join(
        combined_all %>% select(patient_id, siteid)) %>%
      mutate(siteid = "OVERALL")) %>%
  group_by(siteid, patient_id) %>% 
  summarise(
    has_d2830 = any(days_post_enrollment >= 28 & days_post_enrollment <= 30 & has_outcome_value),
    has_d2835 = any(days_post_enrollment >= 28 & days_post_enrollment <= 35 & has_outcome_value)) %>%
  group_by(siteid) %>%
  summarise(
    n = n(),
    n_d2830 = sum(has_d2830),
    pct_d2830 = mean(100 * has_d2830) %>% round(1),
    n_d2835 = sum(has_d2835),
    pct_d2835 = mean(100 * has_d2835) %>% round(1)
  )
```

Compute, for each patient, which is their closest measurement to day 28, where ties are resolved by taking the *later* day:

```{r}
best_time_tbl <-
  patient_time_grid %>%
    # add site information
    left_join(combined_all %>% 
                  select(patient_id, siteid)) %>%
    group_by(siteid, patient_id) %>% 
    filter(has_outcome_value) %>%
    group_by(siteid, patient_id) %>%
    mutate(
        disp_from_28d = days_post_enrollment - 28
    ) %>%
    summarise(
        min_disp_from_28d = disp_from_28d[
            abs(disp_from_28d) == min(abs(disp_from_28d))] %>%
            # Select the measurement *after* d28 if either side
            max(),
        closest_measurement_earlier =
          28 + max(disp_from_28d[disp_from_28d <= 0]) %>%
          (function(x) if (x == -Inf) {NA} else {x}),
        closest_measurement_later   =
          28 + min(disp_from_28d[disp_from_28d >= 0]) %>%
          (function(x) if (x == Inf) {NA} else {x}),
        closest_measurement = 28 + min_disp_from_28d
    )
```

Best times outside d28-30:

```{r}
best_time_tbl %>%
  filter(!(closest_measurement %in% 28:30)) %>% 
  mutate(patient_id = str_replace(patient_id, ".*__", "")) %>% 
  select(-min_disp_from_28d) %>% 
  arrange(siteid, closest_measurement, patient_id) %>%
  knitr::kable()
```

### Baseline missingness

In tabular form:

```{r}
(missingness_table_overall <-
  naniar::miss_var_summary(combined_all)) %>%
  head() %>%
  knitr::kable(digits = 0)
```

```{r}
non_site_vars <- names(combined_all)[names(combined_all) != "siteid"]

(missingness_table_by_site <-
  combined_all %>%
    nest(data = all_of(non_site_vars)) %>%
    mutate(x = map(data, naniar::miss_var_summary)) %>%
    select(-data) %>%
    unnest(x)) %>%
  head() %>%
  knitr::kable(digits = 0)
```

Site-wise, as a wide table:

```{r}
var_categories <- read_tsv(
  "variable_categories.tsv")

missingness_table_by_site_wide <-
  missingness_table_by_site %>%
  select(-n_miss) %>%
  mutate(pct_miss = round(pct_miss)) %>%
  pivot_wider(names_from = siteid, values_from = pct_miss) %>%
    left_join(var_categories)
```

Coloring this:

```{r}
require(gt)
missingness_table_by_site_wide %>%
  left_join(naniar::miss_var_summary(combined_all)) %>%
  group_by(category) %>%
  arrange(desc(pct_miss)) %>%
  mutate(pct_miss = round(pct_miss)) %>%
  gt::gt(rowname_col = "variable", groupname_col = "category") %>%
  data_color(
    columns = unique(combined_all$siteid),
    colors = scales::col_numeric(
      palette = c("lightgrey", "orange"),
      domain = c(0, 100))
    ) %>%
  data_color(
    columns = "pct_miss",
    colors = scales::col_numeric(
      palette = c("lightgrey", "cornflowerblue"),
      domain = c(0, 100)
  ))
```

## Imbalance in covariates

> Before the outcome data are shared with the analysis team, we will examine the distributions of
baseline covariates within and between trials. 

> If there are substantial covariate imbalances (with respect to covariates not included in our
model) between treated and control groups, either within a larger trial or overall, we will consider
including that covariate in our outcome model.

Therefore: need to check imbalance on all the (baseline) covariates (or at least **not just those to include in the model**,  **overall** and **within each trial**).

MatchIt is probably an easy way to do this.

First, to assign treatment groups.

```{r}
distinct(combined_all %>% select(siteid, rand_grp)) %>% knitr::kable()
```

```{r}
(study_arm <-
  read_tsv("../resources/study_arm_tbl.csv"))
```

```{r}
combined <- left_join(
  combined_all %>% filter(rand_grp != "SCREEN FAIL"), # Filter NYU screen fale
  study_arm)

combined %>% select(siteid, rand_grp, treat, azithro) %>% distinct()
```

How many in each arm, by study and overall?

```{r}
combined %>%
  bind_rows(
    combined %>% mutate(siteid = "[OVERALL]")
  ) %>%
  count(siteid, treat)
```

### Add factor versions of variables

```{r}
niaid_levels <- c(
  "1 - death",
  "2 - hosp, mech vent",
  "3 - hosp, vent",
  "4 - hosp, supp ox",
  "5 - hosp, no ox",
  "6 - no hosp, lim",
  "7 - no hosp, no lim")

combined_2 <-
  combined %>%
  mutate(
    covid_scale_baseline_fct = 
      factor(covid_scale_baseline, 
        levels = 1:7,
        labels = niaid_levels,
        ordered = TRUE),
    sex_fct = 
      factor(
        sex,
        levels = 1:2,
        labels = c("male", "female")) %>%
      forcats::fct_explicit_na(),
    race_fct =
      factor(
        race,
        levels = 1:8,
        labels = c(     
          "1 = American Indian/Alaska Native",
          "2 = Asian",
          "3 = Black/African American",
          "4 = Native Hawaiian/Pacific Islander",
          "5 = White",
          "6 = Multiple", 
          "7 = Other/declined", 
          "8 = Unknown/unavailable")),
    ethnic_fct = factor(
      ethnic,
      levels = 0:2,
      labels = c("not hispanic", "hispanic", "unknown")),
    race_simplified_fct =
      factor(race_simplified),
    azithro_fct =
      factor(
        azithro, 
        levels = c(F, T), 
        labels = c("no azithro", "azithro"))
    )

write_rds(combined_2, "output/formatted_baseline_tbl.rds")
```

### Overall imbalance

#### BMI

```{r fig.height=5.5, fig.width=11}
balance_plot_cts <- function(dataframe, covariate) { 
  ggplot(
  dataframe, aes(treat, {{covariate}})) + 
    # Alternatively: beeswarm
  ggbeeswarm::geom_quasirandom(size = .7, groupOnX = TRUE, aes(color = treat)) +
  scale_color_manual(values = c("darkgreen", "darkorange")) +
  facet_grid(~siteid, scales = "free_x", space = "free_x") +
  geom_boxplot(color = "grey40",
               fill = NA,
               alpha = .4, 
               outlier.shape = NA, 
               coef = 0, size = .6) +
  background_grid(major = "y", color.major = "white") +
  theme(axis.text.x = element_text(angle = 90, vjust = .5),
        strip.text.x = element_text(size = 11),
        panel.background = element_rect(fill = "grey95")) + 
  geom_tile(alpha = 0) # deals with blank facets
}

balance_plot_cts(combined %>% 
                   bind_rows(
                     combined %>% mutate(siteid = "[OVERALL]")),
                 bmi)
```

### Age

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
  combined_2 %>% 
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
    mutate(age_5y = age_5y + rnorm(n())),
  age_5y) +
  scale_y_continuous(
    breaks = scales::breaks_width(10))
```

#### Baseline outcome scale

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
  combined_2 %>% 
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
    mutate(covid_scale_baseline = covid_scale_baseline + rnorm(n(), sd = .1)),
  covid_scale_baseline) +
  scale_y_continuous(
    breaks = 2:5,
    labels = (function(b) {niaid_levels[b]}))
```

UMC seems to have a slightly surprising imbalance?

Check the association of this with age (for the Duke cases too), to see if the imputation is likely to be reasonable:

```{r fig.height = 4, fig.width = 6}
ggplot(combined, aes(age_5y, covid_scale_baseline)) +
  geom_jitter(size = .6, width = 1.5, height = .25) +
  geom_vline(aes(xintercept = age_5y), 
             data = combined %>% filter(str_detect(siteid, "Duke")),
             color = "grey80")
```

#### Symptom onset days before enrollment

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
    combined %>% 
        bind_rows(
            combined %>% mutate(siteid = "[OVERALL]")) %>%
      mutate(
        sym_onst_days_bfr_enrdt =
          sym_onst_days_bfr_enrdt + runif(n(), -.3, .3)),
    sym_onst_days_bfr_enrdt)
```

-   New Orleans has an individual with a negative value here. 

-   Duke is missing all of the values (n = 11).

-   The NYU imbalance is perhaps a bit surprising? hard to tell.

What do the marginal distributions of this look like?

```{r}
gg_hist(combined_2, sym_onst_days_bfr_enrdt) +
  facet_rep_grid(treat ~.)
```

Is this associated with baseline COVID status?

```{r fig.height=6, fig.width=5}
combined_2 %>%
  mutate(sym_onst_days_bfr_enrdt = sym_onst_days_bfr_enrdt + rnorm(n(), sd = .3)) %>%
  ggplot(aes(covid_scale_baseline_fct, sym_onst_days_bfr_enrdt)) +
  geom_beeswarm(size = .2) +
  geom_boxplot(alpha = .1, fill = "orange", outlier.shape = NA, width = .4)
```

```{r}
combined_2 %>%
  mutate(
    age_5y = age_5y + rnorm(n(), sd = 1),
    sym_onst_days_bfr_enrdt = sym_onst_days_bfr_enrdt + rnorm(n(), sd = .2)) %>%
  gg_scatter(age_5y, sym_onst_days_bfr_enrdt) +
  geom_smooth(se = FALSE, color = "cornflowerblue", size = .6)
```

#### Date of admission

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
    combined %>% 
        bind_rows(
            combined %>% mutate(siteid = "[OVERALL]")),
    adm_days_before_enrdt)
```

#### Screening days before enrollment

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
    combined %>% 
        bind_rows(
            combined %>% mutate(siteid = "[OVERALL]")),
    scr_days_before_enrdt)
```


#### First dose days after enrollment

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
    combined %>% 
        bind_rows(
            combined %>% mutate(siteid = "[OVERALL]")),
    frstdose_days_after_enrdt)
```

#### Randomization days after enrollment

```{r fig.height=5.5, fig.width=11}
balance_plot_cts(
    combined %>% 
        bind_rows(
            combined %>% mutate(siteid = "[OVERALL]")),
    rand_days_after_enrdt)
```

### Discrete covariates

#### Sex

Try a plot first:

```{r fig.height=9, fig.width=5}
combined_2 %>% 
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  gg_hist(sex_fct) +
  facet_rep_grid(siteid~treat, scale = "free_y") +
  scale_y_continuous(
    breaks = scales::pretty_breaks(2),
    expand = expansion(c(0, .1))
  ) +
  theme(strip.background = element_blank(),
        strip.text.y = element_text(angle = 0))
```


```{r}
combined_2 %>%
  bind_rows(
    combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  group_by(siteid) %>% 
  count(treat, sex_fct) %>%
  ungroup() %>%
  group_by(siteid, treat) %>%
  mutate(pct = round(100 * n/sum(n)), total = sum(n)) %>%
  ungroup() %>%
  pivot_wider(names_from = c(sex_fct), 
              values_from = c(pct))
```

Simplifying to percent female:

```{r}
combined_2 %>%
  bind_rows(
    combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  group_by(siteid) %>% 
  count(treat, sex_fct) %>%
  group_by(siteid, treat) %>%
  mutate(arm_total = sum(n),
         female_pct = round(100*n/arm_total)) %>%
  filter(sex_fct == "female") %>%
  select(-sex_fct, -n) %>%
  ungroup() %>%
  group_by(siteid) %>%
  mutate(site_total = sum(arm_total)) %>%
  select(-arm_total) %>%
  pivot_wider(names_from = treat, values_from = female_pct)
  
```

#### Race and ethnicity

```{r fig.height=9, fig.width=8}
combined_2 %>%
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  gg_hist(race_simplified_fct) +
  facet_rep_grid(siteid~treat, scale = "free_y") +
  scale_y_continuous(
    breaks = scales::pretty_breaks(2),
    expand = expansion(c(0, .1))
  ) +
  theme(strip.background = element_blank(),
        strip.text.y = element_text(angle = 0))
```

```{r fig.height=9, fig.width=5}
combined_2 %>%
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  gg_hist(ethnic_fct) +
  facet_rep_grid(siteid~treat, scale = "free_y") +
  scale_y_continuous(
    breaks = scales::pretty_breaks(2),
    expand = expansion(c(0, .1))
  ) +
  theme(strip.background = element_blank(),
        strip.text.y = element_text(angle = 0))
```

#### Baseline covid scale

```{r fig.height=9, fig.width=7}
combined_2 %>%
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  gg_hist(covid_scale_baseline_fct) +
  facet_rep_grid(siteid~treat, scale = "free_y") +
  scale_y_continuous(
    breaks = scales::pretty_breaks(2),
    expand = expansion(c(0, .1))
  ) +
  theme(strip.background = element_blank(),
        strip.text.y = element_text(angle = 0))
```

#### Azithro

```{r fig.height=9, fig.width=5}
combined_2 %>%
    bind_rows(combined_2 %>% mutate(siteid = "[OVERALL]")) %>%
  gg_hist(factor(azithro)) +
  facet_rep_grid(siteid~treat, scale = "free_y") +
  scale_y_continuous(
    breaks = scales::pretty_breaks(2),
    expand = expansion(c(0, .1))
  ) +
  theme(strip.background = element_blank(),
        strip.text.y = element_text(angle = 0))
```

Corresponding table:

```{r}
combined_2 %>%
  bind_rows(
    combined_2 %>%
      mutate(siteid = "[OVERALL]")) %>%
  count(
    siteid, treat,
    azithro_fct) %>%
  group_by(siteid) %>%
  mutate(site_total = sum(n),
         prop = n/site_total) %>%
  select(-n) %>%
  pivot_wider(
    names_from = c(treat, azithro_fct), 
    values_from = prop) %>%
  ungroup() %>%
  gt(rowname_col = "siteid") %>%
  data_color(
    columns = c("HCQ_no azithro","HCQ_azithro","no_HCQ_no azithro","no_HCQ_azithro"),
    colors = scales::col_numeric(
      palette = c("grey95", "orange"),
      domain = c(0, 1))
    ) %>%
  gt::fmt_percent(c("HCQ_no azithro","HCQ_azithro","no_HCQ_no azithro","no_HCQ_azithro"),
                  decimals = 0)
  
```

## Extreme values of covariates

> If there are individuals with extreme values of
baseline covariates, we will check their data with trial investigators, and potentially exclude them
from the outcome analysis.

[Covered above.]

* * * * * * *
* * * * * * *

## Charlson -- exploration

```{r}
charlson_indicators <- 
  var_categories$variable[
    (var_categories$category == "charlson") &
    (var_categories$variable != "charlson")
  ]
```

```{r}
charlson_matrix <-
  combined_2[charlson_indicators] %>%
  as.matrix()

table(charlson_matrix, useNA = "always")
```

```{r}
charlson_long <-
  combined_2 %>%
  select(siteid, patient_id, all_of(charlson_indicators)) %>%
  pivot_longer(-c(siteid, patient_id))

participant_charlson_1_count <-
  charlson_long %>% group_by(siteid, patient_id) %>%
  summarise(charlson_count = sum(value == 1, na.rm = TRUE))

gg_hist(
  participant_charlson_1_count,
  factor(charlson_count))
```

```{r fig.height=10, fig.width=7}
variable_charlson_1_count <-
  charlson_long %>% 
  group_by(name) %>%
  summarise(charlson_var_count = sum(value == 1, na.rm = TRUE))
  

charlson_indicator_plot_df <-
  charlson_long %>%
  left_join(participant_charlson_1_count) %>%
  left_join(variable_charlson_1_count) %>%
  mutate(
    value = forcats::fct_explicit_na(factor(value)),
    name = factor(
      name,
      levels = (arrange(variable_charlson_1_count, desc(charlson_var_count)))$name,
      ordered = TRUE),
    patient_id = factor(
      patient_id,
      levels = (arrange(participant_charlson_1_count, charlson_count))$patient_id,
      ordered = TRUE))
  
(charlson_indicator_plot <-
  charlson_indicator_plot_df %>%
  ggplot(aes(name, patient_id)) +
  geom_tile(aes(fill = value)) +
  scale_fill_manual(values = c("lightgrey", "darkgreen", "orange", "cornflowerblue", "tomato")) +
  facet_grid(siteid~., scale = "free_y", space = "free_y") +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = .5),
        axis.line.y = element_blank()) +
    ylab("patients sorted by #(comorbidity == 1)"))
  
```

Notice that the weirdly coded Charlson variables are all within variables that are missing from ORCHID. So we arguably don't need to worry about them so much.

Plotting number of conditions against outcome scale:

```{r}
combined_2 %>%
  left_join(participant_charlson_1_count) %>%
  mutate(
    covid_scale_baseline = 
      covid_scale_baseline + rnorm(n(), sd = .1)
    ) %>%
  ggplot(aes(factor(charlson_count), covid_scale_baseline)) +
  ggbeeswarm::geom_quasirandom(size = .7, groupOnX = TRUE) +
  scale_y_continuous(
    breaks = 2:5,
    labels = (function(b) {niaid_levels[b]}))
```

Trying a histogram of the same:

```{r}
combined_2 %>%
  left_join(participant_charlson_1_count) %>%
  mutate(
    covid_scale_baseline = 
      covid_scale_baseline + rnorm(n(), sd = .1)
    ) %>%
  ggplot(aes(factor(charlson_count))) +
  geom_histogram(stat = "count") +
  scale_y_continuous(expand = expansion(c(0, .1))) +
  facet_rep_grid(covid_scale_baseline_fct~.,
                 scale = "free_y")
```

Doesn't appear particularly predictive of baseline status (i.e., baseline status doesn't appear particularly helpful for a score). There's selection on baseline status, which probably makes a difference.

### Alternative Charlson scores

First, pull out the useable variables:

```{r}
charlson_indicators_nonmissing <-
  charlson_indicators[!(charlson_indicators %in% c("aids", "vaping", "hiv", "smoking"))]
```

Table with different options for weights:

```{r}
charlson_weight_tbl <-
  read_tsv("2021-06-14_charlson_weighting.tsv")
```

```{r}
charlson_weight_tbl %>% 
  select(
    variable, missing, 
    freilich_weight, adjusted_log_or_rescaled, log_or_rescaled) %>% 
  pivot_longer(
    cols = c(freilich_weight, adjusted_log_or_rescaled, log_or_rescaled)) %>% 
  filter(!missing) %>% 
  qplot(variable, value, color = name, data = .) + 
  geom_line(aes(group = name)) +
  coord_flip()
```

```{r}
weights_tbl <-
  combined_2 %>% 
  select(patient_id, siteid, 
         all_of(charlson_indicators_nonmissing)) %>% 
  pivot_longer(cols = all_of(charlson_indicators_nonmissing)) %>% 
  rename(variable = name) %>% 
  left_join(charlson_weight_tbl) %>% 
  group_by(siteid, patient_id) %>% 
  summarise(
    freilich_weight = sum(freilich_weight*value), 
    adj_log_or_weight = sum(adjusted_log_or_rescaled*value, na.rm = TRUE), 
    log_or_weight = sum(log_or_rescaled*value, na.rm = TRUE),
    equal_weight = sum(value, na.rm = TRUE)) %>%
  mutate(
    equal_weight_jittered = equal_weight + rnorm(n(), sd = .06))
```

```{r}
plot_grid(
  gg_hist(weights_tbl, freilich_weight),
  gg_hist(weights_tbl, adj_log_or_weight),
  gg_hist(weights_tbl, log_or_weight),
  gg_hist(weights_tbl, equal_weight),
ncol = 1)
```

```{r fig.height=10, fig.width=5}
plot_grid(
  gg_scatter(weights_tbl, equal_weight_jittered, adj_log_or_weight) + theme(aspect.ratio = 1),
  gg_scatter(weights_tbl, equal_weight_jittered, log_or_weight) + theme(aspect.ratio = 1),
  gg_scatter(weights_tbl, equal_weight_jittered, freilich_weight) + theme(aspect.ratio = 1),
ncol = 1)
```

```{r}
# Pairs plot
```

## Considering degrees of freedom in a spline basis

```{r fig.height=11, fig.width=10, warning=FALSE, message=FALSE}
bmi_ns_1df <- splines::ns(combined_2$bmi, df = 1)
plots_1df <-
  list(
    qplot(combined_2$bmi, bmi_ns_1df %*% rnorm(1)),
    qplot(combined_2$bmi, bmi_ns_1df %*% rnorm(1)),
    qplot(combined_2$bmi, bmi_ns_1df %*% rnorm(1)),
    qplot(combined_2$bmi, bmi_ns_1df %*% rnorm(1)))

bmi_ns_2df <- splines::ns(combined_2$bmi, df = 2)
plots_2df <-
  list(
    qplot(combined_2$bmi, bmi_ns_2df %*% rnorm(2)),
    qplot(combined_2$bmi, bmi_ns_2df %*% rnorm(2)),
    qplot(combined_2$bmi, bmi_ns_2df %*% rnorm(2)),
    qplot(combined_2$bmi, bmi_ns_2df %*% rnorm(2)))

bmi_ns_3df <- splines::ns(combined_2$bmi, df = 3)
plots_3df <-
  list(
    qplot(combined_2$bmi, bmi_ns_3df %*% rnorm(3)),
    qplot(combined_2$bmi, bmi_ns_3df %*% rnorm(3)),
    qplot(combined_2$bmi, bmi_ns_3df %*% rnorm(3)),
    qplot(combined_2$bmi, bmi_ns_3df %*% rnorm(3)))

bmi_ns_4df <- splines::ns(combined_2$bmi, df = 4)
plots_4df <-
  list(
    qplot(combined_2$bmi, bmi_ns_4df %*% rnorm(4)),
    qplot(combined_2$bmi, bmi_ns_4df %*% rnorm(4)),
    qplot(combined_2$bmi, bmi_ns_4df %*% rnorm(4)),
    qplot(combined_2$bmi, bmi_ns_4df %*% rnorm(4)))

bmi_ns_5df <- splines::ns(combined_2$bmi, df = 5)
plots_5df <-
  list(
    qplot(combined_2$bmi, bmi_ns_5df %*% rnorm(5)),
    qplot(combined_2$bmi, bmi_ns_5df %*% rnorm(5)),
    qplot(combined_2$bmi, bmi_ns_5df %*% rnorm(5)),
    qplot(combined_2$bmi, bmi_ns_5df %*% rnorm(5)))

plot_grid(
  plotlist = c(plots_1df, plots_2df, plots_3df,
               plots_4df, plots_5df), 
  nrow = 5)
```

## Scatterplot of variables to use in imputation model

```{r fig.height=7, fig.width=7}
combined_2 %>% 
  select(
    covid_scale_baseline,
    sex,
    age_5y,
    bmi,
    sym_onst_days_bfr_enrdt) %>%
  mutate(
    covid_scale_baseline = covid_scale_baseline + rnorm(n(), sd = .15),
    sex = sex + runif(n(), -.1, .1),
    age_5y = age_5y + rnorm(n(), sd = 1),
    sym_onst_days_bfr_enrdt = sym_onst_days_bfr_enrdt + rnorm(n(), sd = .2)
  ) %>%
  pairs(panel = panel.smooth, pch = ".")
```

## Tables for Risk of Bias Assessment

### Baseline covariates by treatment group, and by treatment x study

```{r}
require(table1)
```

Main baseline variables:

```{r}
table1(~ sex_fct + 
         race_fct + ethnic_fct + 
         age_5y +
         bmi +
         covid_scale_baseline_fct +
         sym_onst_days_bfr_enrdt
       | siteid * treat, 
       data = combined_2 %>%
         mutate(
           covid_scale_baseline_fct = 
             factor(covid_scale_baseline_fct, 
                    levels = niaid_levels[2:5])
           ))
```

### Corticosteroids and azithromycin
```{r}
count(combined_2, azithro, on_azithromycin, on_corticosteroids)
```

```{r}
table1(~ on_corticosteroids + on_azithromycin + azithro_arm
       | siteid * treat, 
       data = combined_2 %>%
         mutate_at(
           vars(on_corticosteroids, on_azithromycin),
           as.logical
           ) %>%
         rename(azithro_arm = azithro))
```

Version of azithro table emphasizing compliance or not:

```{r}
table1(~ azithro_use
       | siteid * treat, 
       data = combined_2 %>%
         mutate(
           azithro_use = case_when(
             azithro    & on_azithromycin    ~ "assigned, took",
             (!azithro) & on_azithromycin    ~ "not assigned, took",
             (!azithro) & (!on_azithromycin) ~ "not assigned, did not take",
             azithro    & (!on_azithromycin) ~ "assigned, did not take"
           )
         ))
```

### Charlson variables by treatment group, and by treatment x study


```{r}
charlson_vars_bar_score <-
  var_categories %>%
  filter(category == "charlson",
         variable != "charlson") %>%
  pull(variable)

table1(~ aids + crbr_vasc_dis + mi + 
         cng_hrt_flr + dementia + copd + 
         asthma + hypertension + hiv + 
         tumor + liver_dis + diabetes + 
         smoking + vaping + charlson
       | siteid * treat, 
       data = combined_2 %>%
         mutate_at(
           .vars = charlson_vars_bar_score, 
           .funs = as.logical))
```

### Outcome missingness by treatment group and treatment x study

Using the (28d-35d) definition:

```{r}
left_join(combined_2, patient_time_grid) %>%
  group_by(treat, siteid, patient_id) %>%
  filter((days_post_enrollment >= 28) &
           (days_post_enrollment <= 35)) %>%
  summarise(has_outcome_value = any(has_outcome_value)) %>%
  table1(~ has_outcome_value
       | siteid * treat, 
       data = .)
```

### First and last dose information

```{r fig.height=4, fig.width=16}
combined_2 %>%
  group_by(siteid, treat) %>%
  arrange(lstdose_days_after_enrdt) %>%
  mutate(plot_order = 1:n()) %>%
  ggplot(aes(plot_order)) +
  facet_grid(treat~siteid, scales = "free_x", space = "free_x") +
  geom_linerange(aes(
    x = plot_order, 
    ymin = frstdose_days_after_enrdt, 
    ymax = lstdose_days_after_enrdt),
    color = "grey70") +
  geom_point(aes(x = plot_order, y = frstdose_days_after_enrdt), size = .3) +
  geom_point(aes(x = plot_order, y = lstdose_days_after_enrdt), size = .3) +
  geom_point(aes(x = plot_order, y = died_days_after_enrdt), size = .6, color = "red") +
  xlab("") +
  theme(axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line.y = element_blank(),
        panel.background = element_rect(fill = "grey95"),
        strip.text = element_text(angle = 90, vjust = .5)) +
  background_grid(major = "y", color.major = "white")
```

There appear to be the following kinds of anomaly:

Given the differences in the control arms, we should really care about these *among the HCQ patients*:

```{r}
table1(~ frstdose_days_after_enrdt + 
         lstdose_days_after_enrdt + 
         dose_duration
       | siteid * treat, 
       data = 
         combined_2 %>% 
         mutate(dose_duration = lstdose_days_after_enrdt - frstdose_days_after_enrdt))
```

Inconsistent individuals:

```{r}
combined_2 %>%
  mutate(
    dose_duration = lstdose_days_after_enrdt - frstdose_days_after_enrdt + 1,
    dose_before_enrollment = frstdose_days_after_enrdt < 0,
    negative_dose_duration = dose_duration < 0) %>%
  filter(dose_before_enrollment | negative_dose_duration) %>%
  select(patient_id, rand_grp, frstdose_days_after_enrdt, lstdose_days_after_enrdt) %>%
  knitr::kable()
  
```

## Things to discuss
* * * * *


