---
title: "Simulation"
author: "Leon Di Stefano"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(20200524)

require(tidyverse)
require(cowplot)
require(lemon)
theme_set(theme_cowplot())
theme_update(
  strip.background = element_blank(),
  strip.text.y = element_text(angle = 0),
  strip.text = element_text(face = "bold")
  )

require(ggbeeswarm)
gg_scatter <- function(data, x_var, y_var) {
  qplot({{x_var}}, {{y_var}}, 
        data = data,
        size = I(0.4)) +
    cowplot::theme_cowplot()
}

gg_hist <- function(data, x_var, n_bins = 30) {
  qplot({{x_var}}, bins = I(n_bins), 
        data = data,
        color = I("white")) +
    scale_y_continuous(expand = expansion(0, .05)) +
    cowplot::theme_cowplot()
}

require(naniar)
require(mice)
require(haven)
require(brms)
require(tidybayes)
require(bayesplot)
require(MASS, exclude = "select")
```

## Load and combine the data

Load and combine the data:

```{r}
patients_raw <- read_dta(
  "../data/2021-09-29_HCQ_Analysis_Datasets/HCQ_Analysis_Datasets/8Site_PatientData.dta")

baseline_outcomes_raw <- read_dta(
  "../data/2021-09-29_HCQ_Analysis_Datasets/HCQ_Analysis_Datasets/8Sites_Outcome_BL.dta")

study_arm <- read_tsv(
    "../resources/study_arm_tbl.csv")

combined_raw <- 
  left_join(
    patients_raw %>%
      mutate(patient_id = str_c(siteid, patient_id, sep = "__")) %>%
      # Do we need to do any other patient-level filtering?
      # Yes: see Malathi's document about outcome missingness.
      # In the real analysis this should be its own (involved) step.
      filter(rand_grp != "SCREEN FAIL"), 
    baseline_outcomes_raw %>%
      mutate(siteid = str_replace(siteid, "3-DUke", "3-Duke"),
             patient_id = str_c(siteid, patient_id, sep = "__"))) %>%
  select(-days_post_enrollment) %>%
  rename(covid_scale_baseline = covid_scale) %>% 
  filter(rand_grp != "SCREEN FAIL") %>% # Filter NYU screen fale
  left_join(study_arm)

niaid_levels <- c(
  "1 - death",
  "2 - hosp, mech vent",
  "3 - hosp, vent",
  "4 - hosp, supp ox",
  "5 - hosp, no ox",
  "6 - no hosp, lim",
  "7 - no hosp, no lim")

charlson_all <-
  c("aids", 
    "crbr_vasc_dis", "mi", "cng_hrt_flr", 
    "dementia", "copd", "asthma", "hypertension", 
    "hiv", "tumor", "liver_dis", "diabetes", 
    "smoking", "vaping")

charlson_binary <-
    c("aids", 
    "crbr_vasc_dis", "mi", "cng_hrt_flr", 
    "dementia", "copd", "asthma", "hypertension", 
    "tumor", "liver_dis", "diabetes",
    # See below. Need to check that this is the correct
    # way to collapse these.
    "smoking_binary", "vaping_binary", "hiv_binary")

d_all <-
  combined_raw %>%
  mutate(
    covid_scale_baseline_fct = 
      factor(covid_scale_baseline, 
        levels = 2:5,
        labels = niaid_levels[2:5],
        ordered = TRUE),
    sex_fct = 
      factor(
        sex,
        levels = 1:2,
        labels = c("male", "female")) %>%
      forcats::fct_explicit_na(),
    race_fct =
      factor(
        race,
        levels = 1:8,
        labels = c(     
          "1 = American Indian/Alaska Native",
          "2 = Asian",
          "3 = Black/African American",
          "4 = Native Hawaiian/Pacific Islander",
          "5 = White",
          "6 = Multiple", 
          "7 = Other/declined", 
          "8 = Unknown/unavailable")),
    ethnic_fct = factor(
      ethnic,
      levels = 0:2,
      labels = c("not hispanic", "hispanic", "unknown")),
    race_simplified_fct =
      factor(race_simplified),
    azithro_fct =
      factor(
        azithro, 
        levels = c(F, T), 
        labels = c("no azithro", "azithro")),
    # Create binary smoking, vaping, hiv variables.
    # Need to check that this coding matches that
    # in each study.
    hiv_binary     = (hiv == 0),
    smoking_binary = (smoking == 0),
    vaping_binary  = (vaping == 0)
    ) %>%
  mutate_at(charlson_binary, as.logical)
```

## Baseline imputation

Now to pull out the variables relevant for *baseline* modelling (i.e., to go into `mice`), making sure everything is of the right type for `mice` to use the correct models:

We're going to drop AIDS and HIV (missing from ORCHID):

```{r}
naniar::gg_miss_var(d_all[,charlson_binary])
```

```{r}
d_for_mice <-
  d_all %>%
  select(
    siteid, patient_id,
    sym_onst_days_bfr_enrdt,
    age_5y, sex_fct, race_simplified_fct, ethnic_fct,
    bmi,
    !!charlson_binary,
    covid_scale_baseline_fct
  ) %>%
  # Dropping AIDS and HIV
  select(-hiv_binary, -aids)
```

Check that the correct datatypes are all there—I assume that's what `mice` works off:

```{r}
head(d_for_mice)
```

Inspect the missingness patterns:

```{r}
mice::md.pattern(d_for_mice, rotate.names = TRUE)
```

Run the multiple imputation. Need to ignore `siteid` and `patient_id` variables:

```{r}
mice_fit <- mice::mice(
  d_for_mice %>% select(-siteid, -patient_id),
  seed = 20210524)
```

Have a look at some diagnostics:

We may want to run more iterations to get a better sense of the mixing here:

```{r}
plot(mice_fit)
```


See which methods was used for each variable:

```{r}
mice_fit$method
```

Try some diagnostic plots:

```{r}
# stripplot(mice_fit, bmi~.imp)
# xyplot(mice_fit, bmi ~ sym_onst_days_bfr_enrdt | .imp)
```

These want `mi` to be `numeric` rather than `logical`. The default method is then "predictive mean matching", which always chooses levels from among existing subjects. Perhaps this would be OK:

```{r}
mice_fit_no_logical <- mice::mice(
  d_for_mice %>% 
    select(-siteid, -patient_id) %>%
    mutate_at(vars(matches(charlson_binary)), as.numeric))
```

```{r}
stripplot(mice_fit_no_logical, bmi~.imp)
```

```{r}
xyplot(mice_fit_no_logical, bmi ~ sym_onst_days_bfr_enrdt | .imp)
```

```{r}
stripplot(mice_fit_no_logical)
```

```{r}
densityplot(mice_fit_no_logical)
```

Looks good ... though it looks like there's no default diagnostic plot for baseline COVID scale.

```{r}
xyplot(mice_fit_no_logical, covid_scale_baseline_fct ~ bmi | as.factor(.imp))
```

```{r}
bwplot(mice_fit_no_logical)
```

See
```https://stefvanbuuren.name/fimd/sec-diagnostics.html``` 
for more.

This will need to be checked in some detail (and can be without the outcome data!).

### Pull out a single imputed dataset

Pull out a single imputed dataset to generate outcome values:

```{r}
charlson_binary_no_hiv_aids <-
    c(# "aids", 
    "crbr_vasc_dis", "mi", "cng_hrt_flr", 
    "dementia", "copd", "asthma", "hypertension", 
    "tumor", "liver_dis", "diabetes",
    # See below. Need to check that this is the correct
    # way to collapse these.
    "smoking_binary", "vaping_binary"
    # "hiv_binary"
    )

d_imp_1 <- 
  d_for_mice %>% 
  select(siteid, patient_id) %>%
  bind_cols(
    as_tibble(complete(mice_fit_no_logical))) %>%
  bind_cols(
    d_all %>% select(treat)
  )
```

## Outcome simulation

via a prior-only version of the outcome model.

Compute the required variables

```{r}
charlson_sums <-
  d_imp_1[,charlson_binary_no_hiv_aids] %>%
  rowSums()

d_for_sim <-
  d_imp_1 %>%
  mutate(
    sex_model = ifelse(sex_fct == "female", 1/2, -1/2),
    age_model = (age_5y - 60)/10,
    charlson_sum_model = charlson_sums,
    charlson_sum_fct_model = factor(charlson_sum_model),
    covid_baseline_model = covid_scale_baseline_fct,
    covid_baseline_numeric_model =
      4 - as.numeric(covid_baseline_model),
    # Add a dummy `covid_scale_28d` variable
    covid_scale_28d_dummy =
      sample(1:7, n(), replace = TRUE) %>%
      factor( 
        levels = 1:7,
        labels = niaid_levels,
        ordered = TRUE)
  )
```

Fit a "prior only" outcome model.

First: what are the priors that need to be set?

```{r}
sap_sim_model <-
  brms::bf(
    covid_scale_28d_dummy ~
        treat*(
          sex_model +
          splines::ns(age_model, 3) +
          charlson_sum_model +
          covid_baseline_numeric_model
        ) +
        (1 + treat || siteid) +
        (1 + treat || charlson_sum_fct_model) +
        (1 + treat || covid_baseline_model)
  )

get_prior(
  sap_sim_model,
  family = cumulative(),
  data = d_for_sim)
```

Set priors and draw from a prior-only model:

```{r}
sim_priors <-
  c(
    set_prior(prior = "normal(.3, .5)", class = "b"),
    set_prior(prior = "lognormal(0, .2)", class = "sd"),
    set_prior(prior = "normal(0, 3)", class = "Intercept")
  )

brm_fit_sim <-
  brm(sap_sim_model,
      family = cumulative,
      prior = sim_priors,
      data = d_for_sim,
      sample_prior = "only",
      cores = parallel::detectCores(),
      seed = 20210524)
```

Have a look at some summaries and diagnostics:

```{r}
summary(brm_fit_sim)
```

```{r fig.height=22, fig.width=22}
mcmc_trace (brm_fit_sim)
```

```{r}
y_pp <- posterior_predict(brm_fit_sim, nsamples = 1)

qplot(as.factor(y_pp)) + scale_y_continuous(expand = expansion(c(0, 1)))
```

Take a draw from it, do exploratory plots:

```{r}
d_sim <-
  d_for_sim %>%
  # left_join(d_all) %>%
  mutate(
    covid_scale_28d = 
      factor(y_pp,
             levels = 1:7,
             labels = niaid_levels,
             ordered = TRUE))
```

Check e.g. the association with age:

```{r}
d_sim %>%
  mutate(
    covid_scale_28d = as.numeric(covid_scale_28d) + rnorm(n(), sd = .15),
    age_5y = age_5y + rnorm(n(), sd = .6)
  ) %>%
gg_scatter(
  age_5y, covid_scale_28d
) +
  scale_x_continuous(breaks = scales::breaks_width(5)) +
  background_grid(color.major = "white") +
  scale_y_continuous(breaks = 1:7, labels = niaid_levels) +
  theme(panel.background = element_rect(fill = "grey90"))
  # geom_smooth(se = FALSE) +
  # aes(color = treat)
```

Perhaps it would make more sense to choose nonnegligible model parameters and see if they're recovered.

## Fitting the outcome model

Now fit the intended model [**though only on a single imputed dataset**—the real analysis will pool over all of them]:

```{r}
sap_model <-
  brms::bf(
    covid_scale_28d ~
        treat*(
          sex_model +
          splines::ns(age_model, 3) +
          charlson_sum_model +
          covid_baseline_numeric_model
        ) +
        (1 + treat || siteid) +
        (1 + treat || charlson_sum_fct_model) +
        (1 + treat || covid_baseline_model)
  )

brm_fit <-
  brm(
      sap_model,
      family = cumulative,
      # prior = sim_priors, -- use default vague priors
      data = d_sim,
      cores = parallel::detectCores(),
      seed = 20210524
      # control = list(adapt_delta = 0.99) # apparently helps avoid divergent transitions
      # iter = 4000
      )
```

Have a look at a model summary:

```{r}
summary(brm_fit)
```

The standard deviations are apparently the parameters most likely to show divergent transitions:

```{r fig.height=5, fig.width=12}
bayesplot::mcmc_trace(brm_fit, regex_pars = ".*sd.*")
```

Check a few diagnostics [see [here](https://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup) for more about warnings]:

```{r}
pp_check(brm_fit, type = "bars_grouped", group = c("treat")) +
  scale_x_continuous(breaks = 1:7)
```

## Computing main estimands

Now to compute the various estimands (in way that can be turned into a function?):

### Standardized/marginal effect (for age)

```{r}
d_sim_counterfactual <-
  d_sim %>%
  mutate(
    treat = ifelse(treat == "HCQ", "no_HCQ", "HCQ")
  ) %>%
  select(-covid_scale_28d)

d_sim_counterfactual_ppc <-
  add_predicted_draws(
    d_sim_counterfactual,
    brm_fit, 
    re_formula = NULL # include all group-level effects (inc. study)
    ) %>%
  mutate(covid_scale_28d = .prediction)
```

Let's do **age** first. Let's also begin with the simple plug-in/maximum likelihood estimate within each stratum:

```{r}
mk_age_fct <- function(age) {
  case_when(
    age < 30             ~ "<29",
    age >= 30 & age < 50 ~ "30-49",
    age >= 50 & age < 70 ~ "50-69",
    age >= 70 & age < 79 ~ "70-79",
    age >= 80            ~ "80+"
    )
}

polr_est_confint <-
  function(treat, covid_scale_28d, 
           # Computing the SEs/CIs requires profiling
           # and an extra Hessian computation, which we want to
           # avoid when we don't need it.
           include_se = TRUE) {
    fit <- 
      polr(covid_scale_28d ~ treat,
           Hess = include_se
           )
    
    if (include_se) { 
      ci <- confint(fit)
      summary <- summary(fit)
      
      return(
        tibble(
          logor_hat = summary$coefficients["treatno_HCQ", "Value"],
          logor_95_lower = ci["2.5 %"],
          logor_95_upper = ci["97.5 %"],
          log_se = summary$coefficients["treatno_HCQ", "Std. Error"])) 
    } else {
      return(
        tibble(
          logor_hat = coef(fit),
          ))
    }
  }
  
```

Test this function:

```{r}
polr_est_confint(d_sim$treat, d_sim$covid_scale_28d)
```

Do this within age brackets; plot it:

> Ideally would do this without nesting, so that the function above would actually take two vectors.

```{r}
(plug_in_pointrange_age <-
  d_sim %>%
  mutate(age_fct = mk_age_fct(age_5y)) %>%
  group_by(age_fct) %>%
  summarise(polr_ests = list(polr_est_confint(treat, covid_scale_28d))) %>%
  unnest(polr_ests) %>%
  ggplot(aes(age_fct, logor_hat)) +
  geom_pointrange(
    aes(ymin = logor_95_lower,
        ymax = logor_95_upper),
    color = "grey60",
    position = position_nudge(+.07)
  ) +
  background_grid(major = "y")
 )
```

Note: *not* sharing the cutpoint estimates across the levels.

Now do the posterior predictive version.

First, make a big data frame w/ observed and counterfactual data; this can be used for each of the stratum effects:

```{r}
d_sim_matchedpair_ppc <-
  d_sim_counterfactual_ppc %>%
  ungroup() %>%
  select(-.chain, -.iteration, -.prediction) %>%
  mutate(observed = FALSE) %>%
  bind_rows(
    expand_grid(
      d_sim %>% mutate(.row = 1:n()), 
      .draw = unique(d_sim_counterfactual_ppc$.draw)) %>%
    mutate(observed = TRUE)
  )
```

Now compute, within each age stratum and each draw, the POLR estimand:

```{r}
age_polr_pp <-
  d_sim_matchedpair_ppc %>%
  mutate(age_fct = mk_age_fct(age_5y)) %>%
  group_by(.draw, age_fct) %>%
  summarize(polr_result = 
              list(
                polr_est_confint(
                treat, covid_scale_28d, 
                include_se = FALSE))) %>%
  unnest(polr_result)
```

Model-based intervals on their own:

```{r}
age_polr_pp %>%
  ungroup() %>%
  group_by(age_fct, logor_hat) %>%
  ggplot(aes(age_fct, logor_hat)) +
  stat_pointinterval(.width = c(.95), 
                     .point = median,
                     point_size = 2.5,
                     size = 1)
```

Combining the model-adjusted and plug-in intervals:

```{r}
plug_in_pointrange_age +
  stat_pointinterval(
    .width = c(.95), 
    .point = median,
    point_size = 2.5,
    size = 1,
    data = 
      age_polr_pp %>%
        ungroup() %>%
        group_by(age_fct, logor_hat),
    position = position_nudge(-.07),
    color = "darkgreen")
```

Corresponding exploratory plots:

```{r}
require(colorspace)
niaid_colors <-
  c(
    "1 - death"           = "grey30", 
    "2 - hosp, mech vent" = "grey50", 
    "3 - hosp, vent"      = darken("coral1", .45), 
    "4 - hosp, supp ox"   = darken("coral1", .3), 
    "5 - hosp, no ox"     = "coral1", 
    "6 - no hosp, lim"    = darken("cornflowerblue"), 
    "7 - no hosp, no lim" = "cornflowerblue"
  )

d_sim %>%
  mutate(age_fct = mk_age_fct(age_5y)) %>%
  ggplot(aes(treat, fill = covid_scale_28d)) +
  geom_bar(
    position = "fill",
    color = "white"
    ) +
  facet_grid(.~age_fct) +
  scale_y_continuous(expand = expansion(c(0, .1)), labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  scale_fill_manual(values = niaid_colors)
```

```{r}
d_sim %>%
  mutate(age_fct = mk_age_fct(age_5y)) %>%
  ggplot(aes(treat)) +
  geom_bar(
    color = "white"
    ) +
  facet_grid(.~age_fct) +
  scale_y_continuous(expand = expansion(c(0, .1))) +
  theme(axis.text.x = element_text(angle = 90, vjust = .5)) +
  background_grid(major = "y", color.major = "white", size.major = .6) +
  theme(panel.ontop = TRUE)
```

### Conditional effect (for age)

Create a data grid w/ a range of ages + reference levels of other variables in the model + treatment-control pairs, then draw from the posterior distribution of the response probabilities, marginalizing over study effects:

```{r}
conditional_tbl_age <-
  expand_grid(
    treat = c("HCQ", "no_HCQ"),
    age = 12:90,
    sex_model = 1/2,
    charlson_sum_model = 0,
    covid_baseline_numeric_model = 0,
    charlson_sum_fct_model = "0",
    covid_baseline_model = "5 - hosp, no ox"
  ) %>%
  mutate(
    age_model = (age-60)/10) %>%
  add_fitted_draws(
    brm_fit,
    scale = "response", 
    # No siteid = marginalize over site effects
    re_formula = ~ charlson_sum_fct_model + covid_baseline_model
  )
```

Compute, for each draw and each pair, the effect of interest = relative risk of mechanical ventilation or death:

```{r}
conditional_tbl_age_probs <-
  conditional_tbl_age %>%
  # Filter for mech vent or death
  filter(.category %in% as.character(1:2)) %>%
  group_by(age, .draw, treat) %>%
  summarise(pi_mech_vent_or_death = sum(.value)) 

conditional_tbl_age_rr <-
  conditional_tbl_age_probs %>%
  group_by(age, .draw) %>%
  summarise(
    rr_mech_vent_or_death = 
      pi_mech_vent_or_death[treat == "HCQ"]/
        pi_mech_vent_or_death[treat == "no_HCQ"]
  )
```

```{r}
conditional_tbl_age_probs %>%
  group_by(age, treat) %>%
  # Posterior mean
  summarise(pi_mech_vent_or_death = mean(pi_mech_vent_or_death)) %>%
  ggplot(aes(age, pi_mech_vent_or_death)) +
  geom_line(aes(group = treat, color = treat))
```

```{r}
conditional_tbl_age_rr %>%
  ggplot(aes(age, rr_mech_vent_or_death)) +
  scale_y_log10(labels = scales::label_math()) +
  tidybayes::stat_lineribbon()
```

Why are these bands so extreme (i.e., why is there so much uncertainty)?

```{r}
conditional_tbl_age_probs %>% 
  filter(.draw %in% 1:100) %>% 
  ggplot(aes(age, pi_mech_vent_or_death)) + 
  geom_line(
    aes(group = paste(treat, .draw), 
        color = treat), 
    alpha = .4) + 
  scale_color_manual(values = c("orange", "cornflowerblue"))
```

```{r}
sessionInfo()
```
